HSD17B7 is required for Auditory Function by Regulating Cholesterol Synthesis in Sensory Hair Cells

Version 1

Abstract

An imbalanced cholesterol level is a root cause of various pathologies. However, research on the role of cholesterol homeostasis in hearing physiopathology is still in its infancy. HSD17B7 (17β-Hydroxysteroid dehydrogenase type 7) converts zymosterone to zymosterol in the process of cholesterol synthesis. Here, we found that Hsd17b7 is highly enriched in sensory hair cells of zebrafish and mice. The deficiency of Hsd17b7 led to reduced cholesterol levels in HEI-OC1 cells and hair cells of zebrafish, resulting in compromised MET and auditory function. A heterozygous nonsense variant, c.544G>T (p.E182*) in HSD17B7 was identified in an individual with bilateral profound hearing loss. mRNA of HSD17B7(p. E182*) failed to rescue the impaired MET and auditory function of hsd17b7 mutants. The mutation of HSD17B7 disrupted its interaction with RER1, leading to altered localization and cholesterol distribution. This provides an explanation for the pathogenic effect of the heterozygous mutation observed in the human case and zebrafish. The current study highlights HSD17B7 as a novel gene linked to sensorineural hearing loss.

Introduction

Deafness is a common and highly disabling disease in humans that has not been effectively prevented or controlled. According to a survey report by the World Health Organization, the number of people suffering from deafness worldwide reached 466 million, accounting for more than 5% of the global population. By 2050, nearly 2.5 billion people are projected to have some degree of hearing loss. Deafness has become a major concern for global health. According to the 2016 population survey in China, approximately 70 million people experienced hearing loss. Among children in Jilin, Gansu, Guangdong, and Shanxi provinces of China, the prevalence rate of disabling hearing loss is 0.85%, and it is expected that this prevalence rate will increase year by year [1]. According to pathogenic factors, about 60% of deafness is caused by a genetic defect known as hereditary deafness. Investigating the genetic identification and molecular mechanisms of hereditary deafness would reveal the occurrence and development of the disease, which is the cornerstone of accurate diagnosis and prevention of deafness. Geneticists estimate that the total number of genes responsible for deafness is more than 1000. So far, around 156 non-syndromic deafness genes have been identified [2], and there are still a large number of new deafness genes that need to be identified.

Cholesterol is the most abundant sterol molecule in mammalian cells. It constitutes the cell membrane and is crucial in synthesizing important hormones, facilitating synapse formation, and mediating cell signaling transduction [3–6]. Previous studies have shown that dysregulated intracellular cholesterol homeostasis in auditory cells is involved in auditory development defects [6–8], hereditary hearing loss [9–13], noise-induced hearing loss [14–16], ototoxic hearing loss [17, 18], and age-related hearing loss [19–21]. It has been demonstrated that HSD17B7 converts zymosterone to zymosterol during the cholesterol synthesis process, and is expressed in the brain, heart, eye, and ear [22, 23]. However, the function of HSD17B7 in the auditory system remains unexplored. No pathogenic HSD17B7 mutations have been reported to be associated with hearing loss. Here, we show that HSD17B7 is highly expressed in sensory hair cells in mice and zebrafish. Our findings highlight the critical role of HSD17B7 in the auditory system, positioning it as a potential therapeutic target for hearing loss.

Results

Hsd17b7 expressed in sensory hair cells of zebrafish and mice

To assess the role of HSD17B7 in the auditory system, we first investigated the expression of the hsd17b7 in the developing zebrafish. Single-cell data analysis (accession no. GSE221471) [24] categorized neuromast hair cell (cluster 0), supporting cell (cluster 1), macula hair cell (cluster 2), crista hair cell 1 (cluster 3), mantle cell (cluster 4) and crista hair cell 2 (cluster 5) using the Seurat 4.0.1 platform (Figure 1A). The feature plot and violin plot showed that hsd17b7 was expressed in sensory hair cells, especially in neuromast hair cells (NM HCs) and crista hair cells (Figure 1B, 1C). Then, NM HCs, supporting cells, and mantle cells were ordered along pseudotime using Monocle 3 (Figure 1D, 1E), and a heatmap was generated to visualize the expression dynamics of marker genes in NM HC (Figure 1F). These data indicate that the expression of hsd17b7 gene was gradually increased along the NM HC development trajectory (Figure 1D, 1E). Consistently, whole-mount in situ hybridization (WISH) in zebrafish showed that hsd17b7 was specifically enriched in the NM HCs and crista HCs at 72 hours post-fertilization (hpf) and 96 hpf (Figure 1G). To further explore the localization of hsd17b7 in zebrafish HCs, we generated a hsd17b7-egfp fusion construct driven by a hair cell-specific promoter, myosin 6b [25, 26]. We injected this construct into a fertilized egg and expressed the hsd17b7-egfp fusion in the HCs. Confocal imaging analysis revealed that Hsd17b-EGFP was localized to the cytoplasm in both crista HCs and NM HCs, exhibiting a punctate distribution (Figure 1H, 1I).

High expression of HSD17B7 in zebrafish and mice sensory hair cells

(A) The UMAP analysis of the zebrafish scRNA-seq data. 0, 2, 3, and 5 clusters of cells were annotated as hair cells. 0, neuromast hair cell; 1, supporting cell; 2, macula hair cell; 3, crista hair cell 1; 4, mantle cell; 5, crista hair cell 2. (B) and (C) Feature plot and violin plot of hsd17b7. (D) hsd17b7 expression increased along the pseudotime trajectory of neuromast hair cell formation. (E) Increase in hsd17b7 gene expression levels along the hair cell trajectory. (F) Heatmap of marker genes along the pseudotime trajectory. (G) The expression of the hsd17b7 in 72 hpf and 96 hpf embryos was detected by whole-mount in situ hybridization analysis. Dashed circles indicate the otic vesicle and the neuromast. (H) and (I) Representative images of the crista hair cells and neuromast hair cells in Tg(myo6b: hsd17b7-egfp) at 4 dpf. NM hair cells, neuromast hair cells. Scale bars, 10 μm. (J) The average expression of Hsd17b7 and hair cell marker genes across mouse hair cell types. (K) Immunostaining analysis of HSD17B7 expression in mouse cochlea hair cells at P60. MYO7A was used as a marker for hair cells. The dashed box indicates the magnified region. OHC, outer hair cell; IHC, inner hair cell.

Meanwhile, to further elucidate the expression of HSD17B7 in the mammal auditory system, we analyzed the expression levels of Hsd17b7 in three single-cell data sets of the mouse [27–29], we found that Hsd17b7 were highly expressed in outer hair cells (OHCs), utricles hair cells, and inner hair cells (IHCs), but less in crista hair cells (Figure 1J). Next, we assessed the expression of HSD17B7 in C57BL/6 wild-type mice. Immunostaining analysis revealed that HSD17B7 was expressed in both outer hair cells (OHCs) and inner hair cells (IHCs) (Figure 1K), confirming the scRNA-seq data. These results demonstrate that HSD17B7 is highly expressed in the sensory hair cells of zebrafish and mice.

In addition, phylogenetic and sequence alignment analysis revealed that vertebrate HSD17B7s share a significant similarity in amino acid sequences (Figure S1A). UniProt sequence alignment analysis revealed that Danio rerio Hsd17b7 (NP_001070796.1) shares 58.2% identity with Homo sapiens HSD17B7 (NP_057455.1) and 67.1% identity with Mus musculus HSD17B7 (NP_034606.3), respectively (Figure S1B). These data suggest that HSD17B7 is conserved and expressed in sensory hair cells across vertebrate species.

Hsd17b7 deficiency impaired auditory and MET function in zebrafish

To investigate the Hsd17b7 function in the auditory system, we generated hsd17b7 mutant alleles using CRISPR/Cas9-mediated mutagenesis (Figure S2A, S2B). The target site fragment was PCR-amplified and sequenced (Figure S2C, S2D). Sequence analysis revealed that the mutant alleles included a 4-bp deletion and two 6-bp deletions in exon 3 (Figure S2E). Subsequently, the 4-bp deletion mutant line was chosen for further experiments, which led to a reading frame shift and premature termination of protein translation. The mutant protein is predicted to translate to 96 amino acids, then adds 4 code-shifted residues (Figure S2F).

We first examined the startle response evoked by auditory stimuli to assess the hearing abilities of zebrafish larvae [30]. We found that the movement trajectory, swimming velocity, and distance of the hsd17b7 mutants were significantly decreased compared to those of the controls at 5 dpf (Figure 2A-2C), indicating that depletion of hsd17b7 leads to significant hearing loss. To verify that the hearing loss was specifically caused by hsd17b7 deficiency, we attempted to rescue the impaired auditory function by injecting the hsd17b7 mRNA into hsd17b7 mutants. It was shown that the microinjection of hsd17b7 mRNA successfully restored the compromised movement trajectory, swimming velocity, and distance of the hsd17b7 mutants (Figure 2A-2C), indicating that Hsd17b7 is crucial for the auditory function in zebrafish.

Knockout of hsd17b7 caused compromised MET function and hearing defects in zebrafish

(A) Left: Schematic diagram of the devices used in the startle response assay applied to larvae. Right: Extracted locomotion trajectories from larvae with C-shape motion under a one-time stimulus of 9 dB re.1 ms-2 sound level with 60 Hz tone bursts in control, hsd17b7 KO, and mRNA rescued groups, respectively. Scale bars, 10 mm. (B) and (C) Quantification of the peak velocity and mean distance of movement at 5 dpf larvae under sound stimuli for (A) (n=20). (D) Representative images of neuromast hair cells (HCs, green) in Tg(Brn3C: mGFP) at 4 dpf larvae from control, hsd17b7 KO, and hsd17b7 mRNA rescued groups. Scale bars, 20 μm. (E) Quantification of the number of HCs per neuromast for (D) (n=30). (F) Representative images of neuromast HCs (green) and functional HCs (red) in single neuromast of Tg(Brn3C: mGFP) at 5 dpf larvae from control, hsd17b7 KO, and hsd17b7 mRNA rescued groups, respectively. The white dashed circles indicate NM HCs. Scale bars, 20 μm. (G) Quantification of the FM4-64 relative intensity of HCs per neuromast for (F) (n=27). All quantification data (B, C, E, and G) are presented as the mean ± SD. P values were determined using a one-way ANOVA test followed by Tukey’s multiple comparisons. *P < 0.05; **P < 0.01; ***P < 0.001; ****P < 0.0001; ns, non-significant, p > 0.05.

Furthermore, we designed antisense morpholino oligonucleotides (hsd17b7 Mo) to block the splicing of hsd17b7 RNA at the junction of exon 1/intron 1 to knock down Hsd17b7 to verify the phenotype (Figure S3A, S3B). RT-PCR revealed that natural hsd17b7 mRNA expression was significantly reduced, and mis-spliced mRNA was observed after injection of hsd17b7 Mo (Figure S3C). Additionally, we evaluated the knockdown efficiency and rescue efficiency of hsd17b7 by immunoblotting. As expected, the protein levels of Hsd17b7 decreased significantly after hsd17b7 Mo injection, and mRNA injection effectively rescued the Hsd17b7 protein level (Figure S3D). We subsequently found that the movement trajectory, distance, and swimming velocity of the hsd17b7 morphants were significantly decreased compared to those of controls at 5 dpf (Figure S3E-S3G). Moreover, hsd17b7 mRNA effectively rescued the defective startle response of hsd17b7 morphants (Figure S3E-S3G). These results demonstrate that Hsd17b7 is required for auditory function.

Since Hsd17b7 deficiency impaired auditory function, we investigated whether the mechanotransduction (MET) function of Hsd17b7 mutants is compromised. We generated hsd17b7 mutant alleles in the Tg(Brn3c: mGFP) transgenic line [24, 31], in which the membrane of sensory HCs are labeled with GFP. Confocal microscopy imaging analysis revealed that the average number of HCs in hsd17b7 mutants was decreased compared to that of controls (∼16% reduction) (Figure 2D, 2E). FM4-64 (N-(3-Triethylammoniumpropyl)-4-(6-(4-(Diethylamino) Phenyl) Hexatrienyl) Pyridinium Dibromide) is a fluorescence dye that labels HCs and can rapidly permeate HCs through traversing open mechanosensitive channels, which were commonly used as a proxy for MET function in mature HCs [26, 32, 33]. We found that at 5 dpf, the labeling intensity of FM4-64 in the HCs of the lateral line neuromasts in hsd17b7 mutants was markedly reduced compared with that of controls (∼40% reduction) (Figure 2F, 2G). Meanwhile, we also used the FM4-64 staining assay to examine the MET function of hsd17b7 morphants. The results showed that the FM4-64 intensity in hsd17b7 morphants was also significantly decreased compared to controls at 5 dpf (∼25% reduction) (Figure S4A, S4B). Given that the reduction in HC numbers (∼16%) is much smaller than the decrease in FM4-64 labeling intensity (∼40%), these results suggest that the loss of hsd17b7 function primarily affects the MET function of HCs in zebrafish rather than the number of HCs.

To verify that hsd17b7 is essential for the MET function, we performed a rescue experiment by injecting hsd17b7 mRNA. It was demonstrated that microinjection of the hsd17b7 mRNA successfully restored the FM4-64 intensity and the HC numbers (Figure 2D-2G, S4A, S4B). Collectively, these data demonstrate that hsd17b7 is critical for MET function in zebrafish.

HSD17B7 regulated cholesterol levels in vitro and in vivo

HSD17B7 converts zymosterone to zymosterol, participating in cholesterol biosynthesis (Figure 3A) [23]. Previous research has indicated that abnormally high and low cholesterol levels are detrimental to hearing [7, 34, 35]. To further explore the physiological mechanism underlying how HSD17B7 affects auditory function, we first verified whether HSD17B7 regulates cholesterol. It was shown that HSD17B7 knockdown in HEI-OC1 cells led to a decrease in total cellular cholesterol (Figure 3B, 3C). The minimal cholesterol-binding domain 4 (D4H) cholesterol binding probe derived from the toxin Perfringolysin O (PFO) was widely used to monitor cholesterol at the cytoplasmic leaflet of the plasma membrane (PM) [36, 37]. D4H-mCherry was used to monitor the cholesterol in HEI-OC1 cells. Immunostaining results showed that D4H-mCherry intensity was significantly increased in HEI-OC1 cells overexpressing HSD17B7 compared with control cells (Figure 3D, 3E), suggesting HSD17B7 regulates cholesterol.

HSD17B7 deficiency reduced cholesterol in hair cells

(A) Schematic representation of the cholesterol synthesis pathway. (B) Western blots showing HSD17B7 protein levels in HSD17B7 Knock-down HEI-OC1 cells. Quantification of relative protein levels of HSD17B7 is shown on the right (n = 6). (C) Quantification of the relative cholesterol levels in HSD17B7 knock-down HEI-OC1 cells (n=12). (D) Immunostaining shows cholesterol probe D4H-mCherry in HEI-OC1 cells transfected with pCMV2-Flag and pCMV2-Flag-HSD17B7 full-length plasmids, respectively. Scale bars, 20 μm. (E) Quantification of D4H relative intensity in vitro for (D) (n=16). (F) Schematic diagram of hsd17b7 KO one-cell embryos injection of pmyo6b-D4H-mCherry in Tg(Brn3C: mGFP). (G) Representative images of cholesterol probe D4H-mCherry (red) expression in crista HCs (left) and NM HCs (right) of control or hsd17b7 KO mutant larvae at 4 dpf. Scale bars, 20 μm. (H) Quantification of D4H relative intensity in vivo for (G) (n=30). All quantification data (B, C, E, and H) are presented as the mean ± SD. P values were determined using a two-tailed unpaired Student’s t-test. ***P < 0.001; ****P < 0.0001.

To further assess the functional role of Hsd17b7 in regulating cholesterol in vivo, we injected the construct expressing D4H-mCherry into one-cell-stage embryos of control and hsd17b7 mutant Tg(Brn3c: mGFP) transgenic line to evaluate the cholesterol in zebrafish (Figure 3F). We found that D4H-mCherry was enriched in the hair bundle (Figure 3G), which is consistent with previous reports [36]. Compared with controls, the intensity of D4H-mCherry in crista HCs and NM HCs of hsd17b7 mutants at 4 dpf was significantly reduced (Figure 3G, 3H). These results demonstrate that HSD17B7 regulated cholesterol synthesis in virto and in vivo. Given that abnormal cholesterol homeostasis is detrimental to hearing [7, 34, 35], the deficiency of Hsd17b7 led to compromised MET and auditory function, suggesting that Hsd17b7 regulates auditory function by influencing cholesterol levels in HCs.

A nonsense Variant in HSD17B7 identified in human deafness

Given the established role of Hsd17b7 in auditory function in zebrafish, the human HSD17B7 gene was screened in individuals with hearing loss (HL) who remained undiagnosed after analysis of 201 known HL genes using whole-genome sequencing (WGS)[38]. The proband failed the newborn hearing screening, was enrolled in a special education school during early childhood, and presented at 8 years of age with bilateral profound hearing loss. Physical examination revealed two preauricular appendages anterior to the right ear but no other systemic abnormalities. The proband Ⅱ 1’s parents had normal hearing and no family history of HL but declined genetic testing (Figure 4A, 4B). A heterozygous nonsense variant, c.544G>T (p.E182*) in HSD17B7 (NM_016371.4) was identified and confirmed by Sanger sequencing (Figure 4C). This mutation introduces a premature stop codon at position Glu182 (Figure 4C).

Identification of a novel heterozygote nonsense mutation in HSD17B7

(A) Two-generation family pedigree for the affected individual. The hearing-impaired individual is indicated by a black cycle (female). (B) Pure-tone audiometry audiograms for the hearing-impaired individual at 8 years old. Blue represents the results for the left ear and red for the right ear. Affected individual shows severe-to-profound or profound HI. (C) Sequence of HSD17B7 mRNA in Wild type and Heterozygote. (D) Multiple sequence alignment of 8 different species using TBtools program. The p.Glu182 residue as indicated by a red arrow is evolutionarily conserved from zebrafish to human. (E) The domain structure of human HSD17B7WT and HSD17B7E182*. ED, extracellular domain; TM, transmembrane; CD, cytoplasmic domain. The residue numbers are labeled at right. (F) SWISS-MODEL predicts the three-dimensional protein structure of HSD17B7WT and HSD17B7E182*.

The multiple sequence alignment showed that Glu182 is evolutionarily conserved in eight vertebrates, from zebrafish to humans (Figure 4D), implying that the nonsense mutation may have a potential pathogenetic effect on HI. According to UniProt, the encoded HSD17B7 protein contains two topological domains (extracellular domain: 1-229 aa, cytoplasmic domain: 251-341 aa) and a transmembrane region (230-250 aa). The novel nonsense mutations in the extracellular domain of HSD17B7 (c.544G>T) changed residue 182 from Glu to Ter (p.E182*) and resulted in the production of truncated proteins (Figure 4E). In addition, the three-dimensional protein structure of HSD17B7 is first modeled to predict the effect of the nonsense mutation on the protein using the SWISS-MODEL protein structure database (https://swissmodel.expasy.org) (Figure 4F). This data analysis suggests that this nonsense mutation may have a potential pathogenic effect on hearing loss.

E182* mutation mRNA of HSD17B7 failed to rescue the impaired MET function and auditory behavior of hsd17b7 mutants

To assess the impact of E182* nonsense mutation of HSD17B7 on MET function, we microinjected the human HSD17B7 wild-type or E182* mutant mRNAs separately into the fertilized one-cell stage embryos of hsd17b7 KO mutants in Tg(Brn3c: mGFP) transgenic line. It was shown that wild-type HSD17B7 significantly increased the FM4-64 intensity in NM HCs of hsd17b7 mutants, whereas the HSD17B7 E182* mutant mRNA failed to do so (Figure 5A, 5B), indicating that the E182* nonsense mutation of HSD17B7 has no effect in restoring normal MET function in zebrafish. Furthermore, the startle response assay demonstrated that wild-type HSD17B7 successfully rescued the abnormal auditory behaviors (movement trajectory, distance, and swimming velocity) of hsd17b7 KO mutants, whereas the E182* mutant did not (Figure 5C-5E).

E182* mutation mRNA of HSD17B7 failed to rescue the impaired MET function and auditory behavior of hsd17b7 mutants

(A) Representative images of NM HCs (green) and functional NM HCs (red) in single neuromast of Tg(Brn3C: mGFP) at 5 dpf larvae from control, hsd17b7 KO, KO + E182* mRNA and KO + WT mRNA groups. White dashed circles indicate NM HCs. Scale bars, 20 μm. (B) Quantification of the FM4-64 relative intensity of HCs per neuromast for (A) (n=25). (C) Extracted locomotion trajectories from 5 dpf larvae with C-shape motion under a one-time stimulus of 9 dB re.1 ms-2 sound level with 60 Hz tone bursts in control, hsd17b7 KO, KO + E182* mRNA, and KO + WT mRNA groups, respectively. Scale bars, 10 mm. (D) and (E) Quantification of the mean distance and peak velocity of movement at 5 dpf larvae under sound stimuli for (C) (n=20). All quantification data (B, D, and E) are presented as the mean ± SD. P values were determined using a one-way ANOVA test followed by Tukey’s multiple comparisons. ****P < 0.0001; ns, non-significant, p > 0.05.

Heterozygous E182* mutation of HSD17B7 has a negative effect on the auditory function

To explore the pathogenetic role of HSD17B7 E182* heterozygous mutation in the patient’s hearing loss, we mimicked the proband Ⅱ heterozygote in vitro and vivo. We transiently expressed Flag-tagged HSD17B7 wild-type (HSD17B7WT) or E182* nonsense mutation (HSD17B7E182*) in HEI-OC1 cells to examine their subcellular localization. Since cholesterol synthesis mainly occurs in the endoplasmic reticulum (ER) [39, 40], we supposed that HSD17B7, an enzyme catalyzing cholesterol synthesis, might localize in the ER. Immunostaining revealed that HSD17B7WT co-localized with Calnexin (an ER marker), whereas HSD17B7E182* did not and instead exhibited a spot-like aggregate distribution (Figure 6A). These results were confirmed by matching fluorescence signal intensity profiles (Figure 6B, 6C). Given that the patient was a heterozygote mutant, we suspected that the HSD17B7E182* mutant might impair the subcellular localization of HSD17B7WT. Therefore, we performed immunostaining by co-transfecting HEI-OC1 cells with Flag-HSD17B7WT and HSD17B7E182*-Myc expression constructs. The results showed that the presence of the HSD17B7E182* protein did not alter the subcellular localization of HSD17B7WT (Figure 6D), as confirmed by matching two fluorescence signal intensity profiles (Figure 6E).

Subcellular localization of HSD17B7 WT and E182* mutation

(A) Immunostaining shows the subcellular localization of Flag-HSD17B7WT and Flag-HSD17B7E182* in HEI-OC1 cells. Calnexin was used as a marker for ER. The white arrow indicates the profile position in (B, and C). Scale bars, 20 μm. (B) and (C) The intensity profile shows the localization of HSD17B7WT and HSD17B7E182* with ER, respectively. (D) Immunostaining shows the subcellular localization of HSD17B7WT and HSD17B7E182* in one HEI-OC1 cell. The white arrow indicates the profile position in (E). Scale bars, 20 μm. (E) The intensity profile shows the localization of HSD17B7WT and HSD17B7E182* in one HEI-OC1 cell.

To further investigate the underlying cause of hearing loss in the patient carrying the heterozygous HSD17B7E182* mutation, we injected human wild-type (HSD17B7WT) or mutant (HSD17B7E182*) HSD17B7 mRNAs into one-cell stage embryos of the wild-type in Tg(Brn3c: mGFP) transgenic line to assess whether the overexpression of HSD17B7E182* impairs MET function and auditory behaviors. The FM4-64 assay revealed that the overexpression of HSD17B7WT did not result in a significant difference in fluorescence intensity compared to the control. In contrast, the overexpression of HSD17B7E182* led to a significant reduction of fluorescence intensity (Figure 7A, 7B). In addition, HSD17B7WT overexpression did not alter the startle response compared to the control. In contrast, HSD17B7E182* overexpression resulted in abnormal movement trajectory, along with reduced mean distance and peak velocity in response to auditory stimulation (Figure 7C-7E). These data suggest that overexpression of HSD17B7E182* has a negative effect on normal auditory function.

The heterozygous E182* mutation leads to a negative effect on auditory function

(A) Representative images of neuromast HCs (green) and functional HCs (red) in single neuromast of Tg(Brn3C: mGFP) at 5 dpf larvae from control, HSD17B7E182* mRNA and HSD17B7WT mRNA groups, respectively. White dashed indicate NM HCs. Scale bars, 20 μm. (B) Quantification of the FM4-64 relative intensity of HCs per neuromast for (A) (n=34). (C) Extracted locomotion trajectories from larvae with C-shape motion under a one-time stimulus of 9 dB re.1 ms-2 sound level with 60 Hz tone bursts in control, HSD17B7E182* mRNA and HSD17B7WT mRNA groups, respectively. Scale bars, 10 mm. (D) and (E) Quantification of the mean distance and peak velocity of movement at 5 dpf larvae under sound stimuli for (C) (n=20). (F) Immunostaining shows the distribution of cholesterol probe D4H in HEI-OC1 cells transfected with pCMV2-Flag, pCMV2-Flag-HSD17B7WT, and pCMV2-Flag-HSD17B7E182* plasmids, respectively. The white arrow indicates the profile position in (G, H, and I). Scale bars, 20 μm. (G), (H), and (I) The intensity profile shows the localization of Flag, HSD17B7WT, and HSD17B7E182* with the cholesterol probe D4H, respectively. All quantification data (B, D, and E) are presented as the mean ± SD. P values were determined using a one-way ANOVA test followed by Tukey’s multiple comparisons. *P < 0.05; **P < 0.01; ****P < 0.0001; ns, non-significant, p > 0.05.

Our studies have shown that HSD17B7 regulates cholesterol in vivo and in vitro (Figure 3), and previous research has demonstrated that proper cholesterol distribution in hair cells is essential for maintaining auditory function [36]. Therefore, we hypothesize that the heterozygous HSD17B7E182*mutation may lead to hearing loss by altering cholesterol distribution. Next, we co-transfected HEI-OC1 cells with Flag, Flag-HSD17B7WTor Flag-HSD17B7E182* constructs along with D4H-mCherry (cholesterol probe). Immunostaining results revealed that in cells expressing Flag or Flag-HSD17B7WT, the cholesterol was distributed in the same way around nuclei. In contrast, the nonsense mutation HSD17B7E182* significantly altered the distribution of D4H-mCherry, forming spot-like aggregation (Figure 7F). Furthermore, HSD17B7E182* colocalized with D4H-mCherry, verified by the fluorescence intensity profile (Figure 7G-7I), suggesting that HSD17B7E182* may bind cholesterol and alter its intracellular distribution. Altogether, these results demonstrate that HSD17B7E182* has a negative effect via altering cholesterol distribution in HCs, leading to compromised MET function and auditory behaviors.

Nonsense mutation (p.E182*) of HSD17B7 disrupted the interaction with RER1 and caused the altered localization and cholesterol distribution

To further explore the pathological mechanism of HSD17B7E182*, we transfected HEI-OC1 cells with Flag, Flag-HSD17B7WT, or Flag-HSD17B7E182* constructs and captured the target proteins by co-immunoprecipitation (Figure 8A). We subsequently performed liquid chromatography-tandem mass spectrometry (LC-MS/MS) analysis of the captured proteins (Figure S5). The proteins that interact with HSD17B7WT and HSD17B7E182*, respectively, are shown in the Venn diagram (Figure 8B). We systematically identified the proteins specifically associated with HSD17B7WT and HSD17B7E182*. Consistent with the role of HSD17B7, proteins specifically associated with HSD17B7WT were strongly enriched for cholesterol metabolic process, while proteins specifically associated with HSD17B7 E182* were strongly enriched for chromatin remodeling (Figure 8C). Considering that HSD17B7WT and HSD17B7E182* have different subcellular localizations, we examined the cellular component of HSD17B7WT-specific and HSD17B7E182*-specific interacting proteins. It was found that the HSD17B7WT-specific interacting proteins were enriched in the ER, while the HSD17B7E182*-specific interacting proteins were enriched in the nucleus and cytoplasm (Figure 8D), consistent with the immunostaining results (Figure 6A-6C). Among these ER-associated proteins, RER1 has the top hit (Figure 8E). RER1 is involved in protein retention in ER [41, 42]. We verified the direct interaction between HSD17B7WT and RER1 by an in vitro binding assay. Glutathione S-transferase (GST)-HSD17B7, but not GST, is bound directly to RER1 fused His (RER1-His) (Figure 8F). Next, we co-transfected HEI-OC1 cells with RER1-Myc along with Flag-HSD17B7WT or Flag-HSD17B7E182* constructs. The immunostaining results indicated that RER1 co-localized with HSD17B7WT, whereas it did not co-localize with HSD17B7E182* (Figure 8G), as confirmed by the intensity profiles (Figure 8H, 8I). These data demonstrate that RER1 interacts with HSD17B7WT but not with HSD17B7E182*, suggesting that the localization of HSD17B7WT in the ER is due to its interaction with RER1. In contrast, HSD17B7E182* cannot interact with RER1, resulting in abnormal subcellular localization.

HSD17B7WT binds with Retention in endoplasmic reticulum 1 (RER1)

(A) Flag, Flag-HSD17B7WT, and Flag-HSD17B7E182* were immunoprecipitated from HEI-OC1 cells transfected with pCMV2-Flag, pCMV2-Flag-HSD17B7WT, and pCMV2-Flag-HSD17B7E182* plasmids, respectively. Tubulin and IgG light chains were used as the loading control. Cell lysates were immunoprecipitated with anti-Flag beads. (B) The Venn diagram shows the number of HSD17B7WT interacting proteins (red) and HSD17B7E182* interacting proteins (blue) identified in LC-MS/MS. The number of proteins in each area is marked. (C) Biological Process GO term enrichment analysis for the HSD17B7WT-specific interacting proteins and HSD17B7E182*-specific interacting proteins identified in LC-MS/MS. (D) Cellular Component GO term enrichment analysis for the HSD17B7WT-specific interacting proteins and HSD17B7E182*-specific interacting proteins identified in LC-MS/MS. (E) The top 20 (iBAQ intensity) for the HSD17B7WT-specific interacting proteins in the ER. (F) In vitro binding assays show the interaction between HSD17B7 and RER1. Top: GST and GST-HSD17B7 used for the pull-down assay were stained with Coomassie blue. Bottom: Western blot for RER1-His. (G) Immunostaining shows the localization of RER1-Myc in HEI-OC1 cells transfected with pCMV2-Flag-HSD17B7WT and pCMV2-Flag-HSD17B7E182* plasmids, respectively. The white arrow indicates the profile position in (H, and I). Scale bars, 20 μm. (H) and (I) The intensity profile shows the localization of HSD17B7WT and HSD17B7E182* with RER1, respectively.

During transfection of Flag-HSD17B7WT and Flag-HSD17B7E182* constructs in HEI-OC1 cells, we noticed that the number of positive cells expressing HSD17B7E182* was very low (Figure S6A). Additionally, the IP results also suggested that the nonsense mutation E182* in HSD17B7 led to a significant decrease in protein levels compared to HSD17B7WT (Figure 8A).

To explain the phenomenon, we transfected equal amounts of Flag-HSD17B7WT and Flag-HSD17B7E182* constructs to confirm their protein levels. Immunoblotting revealed that the levels of E182* mutation protein were significantly lower compared to those of the wild-type (Figure S6B), consistent with our aforementioned findings. Next, we measured the mRNA level of HSD17B7WT and HSD17B7E182*. RT-qPCR results showed that the mRNA levels of the E182* mutation were significantly lower compared to those of the wild-type (Figure S6C). Considering that the E182* mutation decreased not only the protein levels but also the mRNA levels, we subsequently examined the effect of the mutation on the half-life periods of mRNA (Figure S6E). Interestingly, the E182* mutation shortened the mRNA half-life periods of HSD17B7 (Figure S6D). Overall, the E182* mutation reduced the stability of HSD17B7 mRNA, resulting in significantly decreased mRNA and protein levels. The abnormally low levels of HSD17B7E182* cannot bind with RER1, preventing proper localization of HSD17B7 to the ER. This, in return, disrupts cholesterol synthesis and distribution, resulting in abnormalities in MET function and auditory behaviors.

Discussion

This study reveals that HSD17B7 is enriched in sensory hair cells in zebrafish and mice, regulating cholesterol levels in hair cells, and is essential for normal mechanotransduction and auditory function. Loss of hsd17b7 markedly reduces cholesterol in hair cells, causing abnormal MET function and hearing behaviors, suggesting that HSD17B7 is a novel candidate gene for sensory hearing loss. This was further supported by a case of sporadic deafness, in which we identified a previously undescribed variant (c.544G>T, p.E182*) in HSD17B7 through whole-exome sequencing and Sanger sequencing analyses. The residual HSD17B7E182* protein alters subcellular localization and disrupts normal cholesterol distribution. The E182 mutation in HSD17B7 reduces the mRNA half-life, decreases mRNA abundance, and significantly lowers protein expression levels. Collectively, our findings suggest that the heterozygous c.544G>T (p.E182*) in the HSD17B7 (NM_016371.4) gene disrupts auditory function through potential pathogenic mechanisms. First, the heterozygous E182* mutation altered the intracellular distribution of cholesterol. Second, the haploinsufficiency resulting from reduced HSD17B7 protein levels disrupts cholesterol biosynthesis. These mechanistic insights highlight the critical role of cholesterol homeostasis in auditory function, providing a molecular basis for the development of potential targeted therapeutic strategies.

HSD17B7 was first identified in the rat prolactin receptor-associated protein [43], which is thought to play a key role in the production of estradiol E2 [44, 45]. Subsequent in vitro experiments demonstrated that HSD17B7 participated in cholesterol synthesis by catalyzing the reduction of zymosterone to zymosterol [46–48]. Both in vivo and in vitro, a deficiency of HSD17B7 leads to decreased cholesterol levels, while overexpression of HSD17B7 results in increased cholesterol levels, consistent with the pathway of cholesterol synthesis. Previous studies have shown that HSD17B7 is expressed in the liver, heart, brain, eye, and ear [22, 23, 49]. Although the role of the HSD17B7 gene in ovarian and breast cancer has been extensively explored in the literature [50–54], the role of the HSD17B7 gene in auditory function remains unclear. Our study highlights that HSD17B7 is essential for normal hearing in vertebrates. This conclusion is supported by the high expression of HSD17B7 in sensory hair cells across vertebrate inner ears, as revealed by the results of whole-mount in situ hybridization and immunostaining in our current study, as well as single-cell RNA sequencing data from our previous publication and those of others [24, 27–29]. Given that loss of hsd17b7 causes abnormal hearing behaviors, our research identifies HSD17B7 as a novel candidate gene for sensory hearing loss.

Cholesterol is a component of biological membranes, and it plays both structural and functional roles in the bio-membranes [55, 56]. Previous studies have shown that abnormally high or low cholesterol levels, or abnormal cholesterol distribution, are detrimental to the auditory system [7, 10, 11, 34–36, 57, 58]. However, they have not explored the effect of cholesterol synthesis by hair cells on this system. Smith–Lemli–Opitz syndrome (SLOS) is a common autosomal recessive disorder caused by mutations in the DHCR7 gene. It is characterized by multiple congenital malformations and intellectual disability resulting from an inborn error of cholesterol synthesis. The syndrome exhibits a wide spectrum of manifestations, ranging from mild behavioral and learning difficulties to perinatal lethality. Additionally, patients commonly present with sensorineural hearing loss; however, auditory function has not been thoroughly investigated [59]. Our study revealed that HSD17B7, an enzyme involved in cholesterol synthesis, is enriched in sensory hair cells and regulates cholesterol levels in these cells, playing a crucial role in maintaining normal auditory function. We found that knocking down or knocking out HSD17B7 resulted in low cholesterol levels, leading to auditory deficiency, whereas overexpressing HSD17B7, which caused elevated cholesterol, did not affect auditory behaviors. Given that the stereocilia links of hair cell kinocilia require cholesterol anchoring, we hypothesize that the hair cell function requires cholesterol to maintain a minimum critical concentration. Cholesterol may stiffen biological membranes [60], a biophysical feature that can be directly involved in MET [61, 62]. Conserved protein-lipid interactions in the TMC1 and TMC2 complex structures of MET-related proteins have been highlighted by cryo-EM [63, 64]. Our research found the truncated variant HSD17B7E182* disturbed cholesterol abundance and distribution in vitro. Considering that the absence of Hsd17b7 or the overexpression of HSD17B7E182* can lead to abnormal MET function in vivo, we hypothesized that HSD17B7 might regulate MET function by affecting the abundance or distribution of cholesterol in hair cells. Our study suggests that cholesterol might regulate hearing function by affecting MET function. Therefore, how cholesterol abnormalities caused by HSD17B7 deficiency or mutations regulate MET function requires further investigation.

Dominant non-syndromic hearing loss is common in patients with hereditary deafness [65–72]. Typically, the dominant allele is inherited from one parent, and de novo mutations should be considered in all cases with a negative family history. Considering that the parents of the proband Ⅱ 1 had normal hearing but were not willing to provide blood samples for sequencing analysis, and combined with the Sanger sequencing analysis of the proband Ⅱ 1, it can be inferred that the proband Ⅱ 1 may have de novo dominant non-syndromic hearing loss. However, this conclusion requires further confirmation through additional clinical cases. Some examples of dominant-negative effects in deafness have been reported, such as KCQN4 and GJB2 [73–76]. Here, we report a case of sporadic deafness with an HSD17B7 (c.544G>T, p.E182*) heterozygous mutation, characterized by bilateral profound deafness with no other clinical features. We found that the mRNA of HSD17B7 (c.544G>T, p.E182*) could not rescue the abnormal auditory function caused by the hsd17b7 KO in a zebrafish. Interestingly, the overexpression of the HSD17B7 mutant mRNA in wild-type zebrafish impaired auditory function, indicating that this nonsense mutation has a negative effect. Our results suggest that the HSD17B7E182* mutant may be due to the altered cholesterol distribution.

In conclusion, we demonstrated that Hsd17b7 is required for auditory function by regulating cholesterol synthesis in sensory hair cells. Hsd17b7 deficiency leads to cholesterol abnormalities, disrupting the MET and auditory function. Additionally, we identified a sporadic nonsense mutation in the HSD17B7 gene in patients with deafness. Our research provides new insights into the critical role of cholesterol in hearing preservation and lays a molecular function for targeted therapeutic strategies.

Materials and Methods

Mouse and zebrafish husbandry

For mouse husbandry, mice in the C57BL/6J background were purchased from the Laboratory Animal Center of Nantong University. Mice were maintained in a barrier facility at 25℃ on a regular 12-hours light and 12-hours dark cycle. The day of vaginal plug observation was embryonic day 0.5 (E0.5), and the day of birth was considered postnatal day 0 (P0). For zebrafish husbandry, zebrafish (Danio rerio) were raised and maintained at 28.5°C. Wild-type AB zebrafish line and the transgenic zebrafish line Tg(Brn3c: mGFP) were used in the study described in our previous study [77]. The embryonic stage is defined as described in the literature [78]. Embryos were collected after natural spawns and moved to E3 medium (5 mM NaCl, 0.17 mM KCl, 0.33 mM CaCl2, and 0.33 mM MgSO4, pH 7.2). After 20 hours post-fertilization (hpf), embryos were moved to a medium containing 0.2 mM phenylthiourea (PTU) to prevent pigmentation for in situ hybridization and imaging. All animal experiments were performed under the guidelines of the Institutional Animal Care and Use Committee of Nantong University.

Cell culture

Mice HEI-OC1 were obtained from Prof. Renjie Chai. The cells were authenticated by short tandem repeat (STR) profiling (Shanghai Biowing Applied Biotechnology Co., Ltd). The cell lines were routinely tested negative for mycoplasma using the MycoBlue Mycoplasma Detection Kit (Vazyme Biotech, #D101-01), and cell aliquots from early passages were used. HEI-OC1 cells were cultured in DMEM (Wisent, #319-005-CL) medium supplemented with 10% FBS (Sigma-Aldrich, #F8318) in an incubator with 10% CO2 at 33℃.

Human HEK293T (sex unknown) cells were obtained from the Shanghai Institute of Biochemistry and Cell Biology (SIBC) and authenticated by short tandem repeat (STR) profilin (Shanghai Biowing Applied Biotechnology Co.,Ltd). The cell lines were routinely tested negative for mycoplasma using the MycoBlue Mycoplasma Detection Kit (Vazyme Biotech, #D101-01), and cell aliquots from early passages were used. Cells were cultured in DMEM (Wisent, #319-005-CL) medium supplemented with 10% FBS (Sigma-Aldrich, #F8318) in an incubator with 5% CO2 at 37℃.

Antibodies

Anti-MYO7A (1:150, Developmental Studies Hybridoma Bank, #138-1); anti-HSD17B7 polyclonal antibody (1:150-1:1000, proteintech, #14854-1-AP); anti-MYC tag polyclonal antibody (1:200, proteintech, #16286-1-AP); anti-His antibody (1:2500, proteintech, # 66005-1-Ig); anti-DYKDDDDK tag mouse monoclonal antibody (1:200-1:5000, proteintech, #66008-4-Ig); anti-calnexin polyclonal antiboy (1:100, proteintech, #10427-2-AP); tubulin monoclonal antibody (1:5000, proteintech, #66031-1-Ig); goat anti-rabbit IgG(H+L) HRP (1:5000, MULTI SCIENCES, #GAR007); goat anti-mouse IgG(H+L) HRP (1:5000, MULTI SCIENCES, #GAM007); FITC goat anti-mouse IgG (H+L) (1:200, ABclonal, #AS001); Cy3 goat anti-rabbit IgG (H+L) (1:200, ABclonal, #AS007); anti-Digoxigenin-AP Fab fragments (1:1000, Roche, #11093274910).

RNA Isolation, Reverse Transcription (RT), and Quantitative Real-time PCR (RT-qPCR)

Total RNAs of all samples were extracted using TRIzol reagent (Invitrogen, #15596026), and genomic DNA contamination was removed by DNase I (Vazyme, #EN401) following the manufacturer’s directions. The RNA yield was determined using NanoDrop ND-2000 (Thermo Fisher Scientific, USA), and integrity was checked on a 1% agarose gel.

The cDNA was synthesized by using HiScript III RT SuperMix (Vazyme, #R323-01) following the manufacturer’s directions. Subsequently, quantitative PCR was performed using the ChamQ SYBR qPCR Master Mix (Vazyme, #Q341-02) with the specified primers (HSD17B7 primer, F: 5’- GACAAGCTTGGATCCATGCGAA-3’, R: 5’- ACCTGGACAATGGTGACCTC-3’; Gapdh primer, F: 5’- CACAGTCAAGGCCGAGAATGGGAAG-3’, R: 5’- GTGGTTCACACCCATCACAAACATG) with a final volume of 20 μL under the following condition: 15 min at 50°C, 5 min at 95°C, and then 40 cycles at 95°C for 15 s and 60°C for 30 s. Relative expression levels for the Hsd17b7 gene were calculated using the 2-△△CT method, normalized to the Gapdh. All reactions were repeated in triplicate for each sample. The results were analyzed using the GraphPad Prism software (version 9.4.0).

Whole-mount in situ hybridization and image acquisition

The whole-mount in situ hybridization (WISH) of zebrafish was performed according to the following standard procedures. A 403 bp cDNA fragment of zebrafish hsd17b7 gene was amplified via PCR using designed primers (F: 5’-GACGTCCTCCAGTAATGCCC-3’, R: 5’- CATCTTGCTTGGTCGGGTGT-3’) and was cloned into the pGEM-T-easy vector. After linearization of the pGEM-T-easy vector inserting the hsd17b7 fragment, the DIG RNA Labeling Kit (SP6) (Roche, #11175025910) was used to prepare digoxigenin-labeled hsd17b7 antisense mRNA probes through transcription in vitro. Subsequently, embryos at different developmental stages were hybridized with an hsd17b7 mRNA probe overnight after a series of treatments, including fixation in 4% paraformaldehyde in PBS, digestion in proteinase K (Roche, #3115879001), and incubation with a pre-hybridized mix. Finally, the alkaline phosphatase (AP)-conjugated antibody against digoxigenin (Roche, #11093274910) and the AP-substrate NBT/BCIP solution (Roche, #11681451001) were used to detect the hsd17b7 expression. The WISH images of zebrafish were acquired using a stereomicroscope (Olympus, MVX10, Japan).

Plasmids construction

The HSD17B7WT and RER1 full-length cDNAs were PCR amplified from the HEK293T cDNA and cloned into the pGEX-TEV and pET-23b vector, respectively. HSD17B7E182* was generated by site-directed mutagenesis. For cell transfection, the cDNA of HSD17B7WT and HSD17B7E182* were subcloned into the pCMV2-Flag vector, respectively. The cDNA of RER1 and HSD17B7E182* were subcloned into the pcDNA3.1-Myc-His A vector, respectively. eGFP was subcloned into pCMV2-Flag-HSD17B7WT and pCMV2-Flag-HSD17B7E182*, respectively. pCS2-D4H-mCherry was obtained from G. Peng’s lab. For plasmid injection of zebrafish, myosin 6b promoter was subcloned into the p-mTol2 vector, followed by D4H-mCherry was subcloned into the p-mTol2 vector. The cDNAs encoding the human protein of HSD17B7WT-eGFP and HSD17B7E182*-eGFP were inserted into the p-mTol2-myo6b vector, then linking P2A-D4H-mCherry by overlap extension. For mRNA injection of zebrafish, the cDNAs encoding the human protein of HSD17B7WT and HSD17B7E182* were inserted into the pCS2+ and likewise, the cDNAs encoding the zebrafish full-length protein of Hsd17b7 were inserted into the pCS2+, which were linearized using NotI and then transcribed in vitro to generate mRNA. All constructs were verified by DNA sequencing.

Cell transfection

HEI-OC1 cells were transfected with plasmids using an X-treme GENE HP DNA transfection reagent (Roche, #6366236001) according to the manufacturer’s protocol. After 36 hours or 48 hours transfection, the cells were fixed or collected and subjected to further analyses, including western blots, Co-IP, IF, or RT-qPCR.

HEI-OC1 cells were transfected with small interfering RNA. The mouse Hsd17b7 siRNA oligos (F: 5’-GGAGGUGUUUGAAACCAAUTT-3’, R: 5’- AUUGGUUUCAAACACCUCCTT-3’) were synthesized from GenePharma (Shanghai, China). Cells were transfected with siRNA oligos to knock down mHSD7B7 using Lipofectamine RNAiMAX (Invitrogen, #13778-150) according to the manufacturer’s protocols. Approximately 1×105 cells per well were placed in six-well plates. When the cell density reached 70-80%, 4 μL of control siRNA or mHsd17b7 siRNA were added to 150 μL of opti-MEM (Gibco, #31985070), then 5 μL Lipofectamine RANiMAX was added, incubated at room temperature for 5 min, then added them to the cell culture medium. After 48 hours transfection, the transfected cells were harvested for western blots or cholesterol assay.

mRNA stability assay

HEI-OC1 cells were transfected with plasmids using an X-treme GENE HP DNA transfection reagent. After 24 hours, the cells were treated with 10 µM actinomycin D (MCE, #HY-17559) for 0, 2, 4, 6 and 8 hours. Then collected cells and isolated RNA for RT-qPCR [79].

Immunostaining and image acquisition

Cultured cells grown on coverslips, after 36 hours transfection, cells were fixed with 4% paraformaldehyde in PBS for 30 min at room temperature and washed 3 times with PBS, then followed by permeabilized in 0.3% PBST (0.3% Triton X-100 in PBS) for 20 min. After three times washes with PBS, the cells were blocked with 10% FBS in PBS, followed by incubation with anti-primary antibodies overnight at 4°C, and then incubated with fluorescent dye-labelled secondary antibodies for 2 hours at room temperature before mounting using the mounting solution with DAPI (1:500, SouthernBiotech, #011-20).

The dissected mouse cochleae were fixed with 4% paraformaldehyde in PBS for 1 hours at room temperature. After 3 times washes with 0.1% PBST, the cochleae were briefly blocked with a blocking medium (PBS containing 1% Triton X-100 and 10% heat-inactivated donkey serum, pH 7.2) for 1 hours at room temperature, followed by incubation with the anti-primary antibody overnight at 4°C. Then the samples were washed in 0.1% PBST and incubated with the fluorescent dye-labelled secondary antibodies for 2 hours at room temperature before mounting using the mounting solution with DAPI (1:500, SouthernBiotech, #011-20).

The immunostaining images of stained HEI-OC1 cells and cochlea were acquired by a Nikon confocal microscope with NIS-Elements software. For zebrafish live imaging, the larvae were anesthetized with tricaine MS-222 (Sigma, #A5040) and mounted in 0.6% low-melting agarose (Invitrogen, #16520050) with a lateral view. All the reconstructed three-dimensional images and adjusted contrast were used with Imaris X64 software (version 9.0.1).

FM4-64 labeling

To investigate the basal activity of neuromast hair cells, the FM4-64 (1:500, Invitrogen, #T13320) vital dye was used to specifically label functional HCs in the neuromasts. The staining procedures were carried out as described previously [25, 26, 32, 33]. The free-swimming larvae were incubated in 3 μM FM4-64 vital dye for 15 s at room temperature in the dark. Afterward, the fish were rinsed 3 times using a PTU medium and imaged. The images of stained hair cells were acquired by a Nikon confocal microscope with NIS-Elements software.

Measurement of total cholesterol

HEI-OC1 cells were homogenized in chloroform/methanol (2:1) to extract lipids and then centrifuged at 20,000 × g for 10 min. The organic phase was harvested and dried using nitrogen flow. Total cellular cholesterol was quantified using an Amplex Red cholesterol assay kit (Invitrogen, #A12216) through a multilabel reader (PerkinElmer, USA) according to the manufacturer’s instructions.

Morpholino and mRNA injections

For inhibiting the expression of hsd17b7, hsd17b7-specific splicing-blocking morpholinos were designed and procured from Gene Tools, Inc. and the precise sequence was (5’- TGCAAACAGGTAACAAAACTGTGTG-3’). The powder of morpholino was dissolved and diluted in RNase-free water to obtain the work solution with a 0.3 mM final concentration for subsequent operations. About 2 nL dose of morpholino work solution was microinjected into zebrafish embryos at the one-cell stage. To examine the morpholino efficiency, embryos injected with morpholino were collected to extract RNA which then was subjected to reverse transcript cDNA. The designed primers franking on exon 1 and exon 2 (F: 5’- TACACAGGCAAACGTTAGAAGC-3’, R: 5’-CTTGAACTCCTCTGCACCCTT-3’) were used to amplify the fragment containing the mis-splicing target site which was located at the connection of exon 1 and intron 1. For rescue experiments, exogenous mRNA was first synthesized by transcription in vitro. Briefly, the designed primers (zebrafish hsd17b7 mRNA primer, F: 5’-CGCGGATCCATGAAGAAAGTAGTTTTGGT-3’, R: 5’- CCGGAATTCTCACATTCCATTTCTTTCTT-3’; human HSD17B7WT mRNA primer, F: 5’- CGGGATCCATGCGAAAGGTGGTTTTGATC-3’, R: 5’- GCTCTAGATTATAGGCATGAGCCACTGA-3’; human HSD17B7E182* mRNA primer, F: 5’-CGGGATCCATGCGAAAGGTGGTTTTGATC-3’, R: 5’-GCTCTAGACTAGAGGCTGAAATTAGATT-3’) were used to amplify target DNA containing the coding sequence. Then, the pCS2 vector inserted into the amplified fragment was linearized as a template for the transcription of mRNA utilizing the SP6 mMESSAGE mMACHINE Kit (Invitrogen, #AM1340). After purifying using the RNeasy Mini Kit (Qiagen, #74104), mRNA with a 70 ng/μL concentration was co-injected into one-cell stage embryos with morpholino for rescue experiments.

sgRNA/ Cas9 mRNA synthesis and injection

sgRNA was designed against the hsd17b7 gene (ENSDARG00000088140) using the CRISPR design tool (https://www.crisprscan.org). The sgRNA specifically targeting exon 3 of hsd17b7 (5’-GGGAATCATGCCTAATCCCA-3’) was first synthesized as follows. The GenCrispr sgRNA synthesis kit is used to generate a gRNA DNA template with a T7 promoter and synthesize gRNA upon in-vitro transcription (Genscript Biotech Co.Ltd., Nanjing, China). The pXT7-zCas9 plasmid was linearized using XbaI and then transcribed in vitro to generate Cas9 mRNA using the mMESSAGE mMACHINE T7 kit (Invitrogen, #AM1344). At the one-cell stage, embryos were injected with a 2-3 µL solution containing 300 ng/µL Cas9 mRNA and 150 ng/µL sgRNA. The injected embryos were then grown in an E3 medium. At 24 hpf, genomic DNA was extracted from injected embryos, and potential CRISPR-induced mutations were identified by PCR. Primers used for identification were designed around the hsd17b7 sgRNA target sites (F: 5’-AAAAACTTATTTTATTCCAGCCCAA-3’, R: 5’- TTTCCATGCAGCACTATCAAACAATT-3’).

Startle Response Test

The acoustic startle reflex was performed as described previously [30, 80]. Briefly, A plastic plate attached to a mini vibrator was used to place 20 normal larvae at 5 days post-fertilization (dpf), while an infrared digital video tracking system was used to monitor their swimming behavior. The 60 Hz tone bursts with two different sound levels of 9 dB re. m s−2 were applied to the amplifier to drive the vibrator. The acoustic vibration stimuli lasting for 30 ms with a 180 s inter-stimulus interval were set and applied. Each sound vibration stimulus level was repeated 20 times and the locomotion behavior of the larvae with C-shape motion to this stimulus was recorded. Finally, the movement typical parameters of mean distance and peak velocity were analyzed to assess the startle response of larvae to sound vibration stimuli.

In vitro binding assay

GST and GST-HSD17B7WT fusion proteins were immobilized to GSH resins were incubated with purified His-RER1 proteins overnight at 4℃ and extensively washed with washing buffer (0.3% PBST). Bound proteins were separated using SDS-PAGE and visualized with Coomassie Blue staining and Western blots.

Western blots

Frozen cells were lysed in RIPA buffer (50 mM Tris-HCl, 150 mM NaCl, 1 mM EDTA, 1% sodium deoxycholate, 1% w/v protein inhibitor (Roche, #4693132001)) for 30 min and centrifuged at 12,000 g for 15 min at 4℃. Protein concentrations were determined using a BCA protein assay kit (Thermo Fisher, #23227).

The proteins were separated on SDS-PAGE gels and subsequently transferred to a PVDF membrane (Millipore, USA). The membranes were blocked in blocking buffer (1 × TBS containing 0.5% milk and 0.5% Tween 20) for about 45 min at room temperature, then incubated overnight with primary antibodies in TBS containing 4% BSA, 1% Tween20, and 0.05% NaN3. Next day, the PVDF membranes were incubated with the appropriate HRP-conjugated secondary goat antibodies for 2 hours at room temperature after extensive washing with a blocking buffer. The blots were detected using ECL TM Western Blotting Detection Reagents (GE, #RPN2106), and acquired images were analyzed with ImageJ software (version 1.8.0).

Co-IP, Mass spectrum, and Gene Ontology (GO) analysis

Cell lysates from Flag, Flag-HSD17B7WT, and Flag-HSD17B7E182* transfected cells were incubated with 30 μL of beads conjugated to anti-Flag antibodies (Smart-Lifesciences, #SA042001) respectively overnight at 4℃. The next day, Flag beads were washed with lysis buffer 3 times, then the immune complexes were eluted with 2 × SDS sample buffer and subjected to three samples for SDS-PAGE and mass spectrum analysis.

For mass spectrum, the immunoprecipitated HSD17B7WT and HSD17B7E182*-associated proteins were washed and dissolved in ammonium bicarbonate buffer (25 mM, pH 8.0), followed by trypsin digestion at 37℃ for 16 hours. The nano-LC-MS/MS experiments were performed with LTQ-Orbitrap MS (Thermo Fisher, USA) equipped with a Nano electrospray ion source.

For Venn diagram and GO analysis, microarray analysis identified HSD17B7WT and HSD17B7E182*-specific associated proteins by mass spectrum (peptide ≥ 1, unique peptide ≥ 1, P value < 0.05). Cellular component GO term enrichment analysis were performed using the web-based DAVID software [81]. Ontology networks were further investigated and visualized using R (version 4.4.1).

ScRNA-seq and data processing

Zebrafish scRNA-seq data was downloaded from NCBI Gene Expression Omnibus (GSE221471) [24]. The basic procedure for single-cell integrated data analysis was carried out in the Seurat 4.0.1 platform [82]. We selected five cell populations for re-analysis based on previous reports, including neuromast hair cells (cluster 0), supporting cells (cluster 1), macula hair cell (cluster 2), cirsta hair cell 1 (cluster 3), mantle cells (cluster 4), and crista hair cell 2 (cluster 5). Gene ontology (GO) and KEGG pathway enrichment analysis were performed using the R package clusterProfiler with the marker genes identified by Seurat V4.0.1 as the input, and the results were visualized using the enrich plot method [83].

To construct single cell trajectories, Monocle 3 (version 1.0.1) was used to cluster cells and reduce the dimensionality of gene expression matrices using UMAP. Monocle 3 package documentation was downloaded from GitHub. Subsets of cells that lead to the hair cell lineage were chosen using the “choose_cells” function for downstream analysis. The cells were ordered in pseudotime according to their development progress using the “order_cells” function and re-clustered in Monocle 3 for pseudotime and differential expression analysis. Cell paths were predicted by the “learn_graph” function of Monocle 3. The clusters identified as mantle cells furthest from the hair cells were chosen as the ‘‘roots’’ of the trajectory. Hair cells (cluster 0), supporting cells (cluster 1), and mantle cells (cluster 2) paths were selected separately (choose_cells) to plot their marker expression along pseudotime (plot_genes_in_pseudotime). Mouse scRNA-seq data was downloaded from previous articles (GSE71982, (GSE168901, GSE202920) [27–29]. The basic procedure for single-cell integrated data analysis and correction of batch effects were carried out in the Seurat 4.0.1 platform [82]. We selected four cell populations for re-analysis, including inner hair cells, outer hair cells, utricle hair cells and crista hair cells.

Whole-exome sequencing and Sanger sequencing analyses

Whole-exome sequencing was performed in the proband Ⅱ 1. Genomic DNA samples were extracted from whole blood samples of the proband Ⅱ 1 (Figure 4A).

The exomes and flanking intronic regions from whole blood DNA samples were captured by Agilent SureSelect Human All Exon Kit (Agilent Technologies, USA). The captured DNA was sequenced on Illumina HiSeq 4000 sequencing platform (Illumina, USA). Bioinformatics were aligned to NCBI build37/hg19 assembly using the BWA (version 0.7.12) software. Each sample was covered to an average sequencing depth of at least 100×. SNPs and indels were identified using GATK HaplotypeCaller software. Candidate pathogenic variants were defined as nonsense, missense, splice-site, and indel variants with allele frequencies of 0.001 or less in public variant databases dbSNP, 1000 Genomes, ESP6500, nci60, GnomAD and in disease database of COSMIC, Clinvar, OMIM, GWAS. Genotypes distributed in every 0.3 cM of genomic region were chosen for calculation of the logarithm of odds (LOD) scores using the Merlin v. 1.1.2 parametric linkage analysis package. The proband Ⅱ 1 was genotyped by Sanger sequencing analyses with designed primers (F: 5’-GTACTCTGATTGGTGACGGGTGAG-3’, R: 5’- GACAGTCATAGTTCATAGTTTATT-3’)

Quantification and statistical analysis

GraphPad Prism 9 (version 9.4.0) supported the whole statistical analysis. All data were repeated more than three times and presented as mean ± SD. An unpaired two-tailed student’s t-test was performed for two-group comparisons, while multiple comparisons were illustrated using one-way ANOVA. The p-value less than 0.05 (p ≤ 0.05) was considered as significantly different. p ≤ 0.05, p ≤ 0.01, p ≤ 0.001, and p ≤ 0.0001 were symbolized with “*, **, ***, and ****”, respectively, and “ns” represented no significance, p > 0.05.

eLife Assessment

This study provides valuable data on the role of Hsd17b7, a gene involved in cholesterol biosynthesis, as a potential regulator of mechanosensory hair cell function. The authors used both zebrafish and the HEI cell line to examine the effects of deletion of Hsd17b7 on hair cell function and survival. While the results do show a reduction in hair cells in the lateral line neuromasts of Hsd17b7 mutant fish, the reduction was limited. The findings are considered incomplete, with additional experiments required to confirm the conclusions.

Reviewer #1 (Public review):

Summary:

This study identifies HSD17B7 as a cholesterol biosynthesis gene enriched in sensory hair cells, with demonstrated importance for auditory behavior and potential involvement in mechanotransduction. Using zebrafish knockdown and rescue experiments, the authors show that loss of hsd17b7 reduces cholesterol levels and impairs hearing behavior. They also report a heterozygous nonsense variant in a patient with hearing loss. The gene mutation has a complex and somewhat inconsistent phenotype, appearing to mislocalize, reduce mRNA and protein levels, and alter cholesterol distribution, supporting HSD17B7 as a potential deafness gene.

While the study presents an interesting candidate and highlights an underexplored role for cholesterol in hair cell function, several important claims are insufficiently supported, and the mechanistic interpretations remain somewhat murky.

Strengths:

(1) HSD17B7 is a new candidate deafness gene with plausible biological relevance.

(2) Cross-species RNAseq convincingly shows hair-cell enrichment.

(3) Lipid metabolism, particularly cholesterol homeostasis, is an emerging area of interest in auditory function.

(4) The connection between cholesterol levels and MET is potentially impactful and, if substantiated, would represent a significant advance.

Weaknesses:

(1) The pathogenic mechanism of the E182STOP variant is unclear: The mutant protein presumably does not affect WT protein localization, arguing against a dominant-negative effect. Yet, overexpression of HSD17B7-E182* alone causes toxicity in zebrafish, and it binds and mislocalizes cholesterol in HEI-OC1 cells, suggesting some gain-of-function or toxic effect. In addition, the mRNA of the variant has a low expression level, suggesting nonsense-mediated decay. This complexity and inconsistency need clearer explanation.

(2) The link to human deafness is based on a single heterozygous patient with no syndromic features. Given that nearly all known cholesterol metabolism disorders are syndromic, this raises concerns about causality or specificity. The term "novel deafness gene" is premature without additional cases or segregation data.

(3) The localization of HSD17B7 should be clarified better: In HEI-OC1 cells, HSD17B7 localizes to the ER, as expected. In mouse hair cells, the staining pattern is cytosolic and almost perfectly overlaps with the hair cell marker used, Myo7a. This needs to be discussed. Without KO tissue, HSD17B7 antibody specificity remains uncertain.

Reviewer #2 (Public review):

A summary of what the authors were trying to achieve.

The authors aim to determine whether the gene Hsb17b7 is essential for hair cell function and, if so, to elucidate the underlying mechanism, specifically the HSB17B7 metabolic role in cholesterol biogenesis. They use animal, tissue, or data from zebrafish, mouse, and human patients.

Strengths:

(1) This is the first study of Hsb17b7 in the zebrafish (a previous report identified this gene as a hair cell marker in the mouse utricle).

(2) The authors demonstrate that Hsb17b7 is expressed in hair cells of zebrafish and the mouse cochlea.

(3) In zebrafish larvae, a likely KO of the Hsb17b7 gene causes a mild phenotype in an acoustic/vibrational assay, which also involves a motor response.

(4) In zebrafish larvae, a likely KO of the Hsb17b7 gene causes a mild reduction in lateral line neuromast hair cell number and a mild decrease in the overall mechanotransduction activity of hair cells, assayed with a fluorescent dye entering the mechanotransduction channels.

(5) When HSB17B7 is overexpressed in a cell line, it goes to the ER, and an increase in Cholesterol cytoplasmic puncta is detected. Instead, when a truncated version of HSB17B7 is overexpressed, HSB17B7 forms aggregates that co-localize with cholesterol.

(6) It seems that the level of cholesterol in crista and neuromast hair cells decreases when Hsb17b7 is defective (but see comment below).

Weakness:

(1) The statement that HSD17B7 is "highly" expressed in sensory hair cells in mice and zebrafish seems incorrect for zebrafish:

(a) The data do not support the notion that HSB17B7 is "highly expressed" in zebrafish. Compared to other genes (TMC1, TMIE, and others), the HSB17B7 level of expression in neuromast hair cells is low (Figure 1F), and by extension (Figure 1C), also in all hair cells. This interpretation is in line with the weak detection of an mRNA signal by ISH (Figure 1G I"). On this note, the staining reported in I" does not seem to label the cytoplasm of neuromast hair cells. An antisense probe control, along with a positive control (such as TMC1 or another), is necessary to interpret the ISH signal in the neuromast.

(b) However, this is correct for mouse cochlear hair cells, based on single-cell RNA-seq published databases and immunostaining performed in the study. However, the specificity of the anti-HSD17B7 antibody used in the study (in immunostaining and western blot) is not demonstrated. Additionally, it stains some supporting cells or nerve terminals. Was that expression expected?

(2) A previous report showed that HSD17B7 is expressed in mouse vestibular hair cells by single-cell RNAseq and immunostaining in mice, but it is not cited:

Spatiotemporal dynamics of inner ear sensory and non-sensory cells revealed by single-cell transcriptomics.

Jan TA, Eltawil Y, Ling AH, Chen L, Ellwanger DC, Heller S, Cheng AG.

Cell Rep. 2021 Jul 13;36(2):109358. doi: 10.1016/j.celrep.2021.109358.

(3) Overexpressed HSD17B7-EGFP C-terminal fusion in zebrafish hair cells shows a punctiform signal in the soma but apparently does not stain the hair bundles. One limitation is the consequence of the C-terminal EGFP fusion to HSD17B7 on its function, which is not discussed.

(4) A mutant Zebrafish CRISPR was generated, leading to a truncation after the first 96 aa out of the 340 aa total. It is unclear why the gene editing was not done closer to the ATG. This allele may conserve some function, which is not discussed.

(5) The hsd17b7 mutant allele has a slightly reduced number of genetically labeled hair cells (quantified as a 16% reduction, estimated at 1-2 HC of the 9 HC present per neuromast). On a note, it is unclear what criteria were used to select HC in the picture. Some Brn3C:mGFP positive cells are apparently not included in the quantifications (Figure 2F, Figure 5A).

(6) The authors used FM4-64 staining to evaluate the hair cell mechanotransduction activity indirectly. They found a 40% reduction in labeling intensity in the HCs of the lateral line neuromast. Because the reduction of hair cell number (16%) is inferior to the reduction of FM4-64 staining, the authors argue that it indicates that the defect is primarily affecting the mechanotransduction function rather than the number of HCs. This argument is insufficient. Indeed, a scenario could be that some HC cells died and have been eliminated, while others are also engaged in this path and no longer perform the MET function. The numbers would then match. If single-cell staining can be resolved, one could determine the FM4-64 intensity per cell. It would also be informative to evaluate the potential occurrence of cell death in this mutant. On another note, the current quantification of the FM4-64 fluorescence intensity and its normalization are not described in the methods. More importantly, an independent and more direct experimental assay is needed to confirm this point. For example, using a GCaMP6-T2A-RFP allele for Ca2+ imaging and signal normalization.

(7) The authors used an acoustic startle response to elicit a behavioral response from the larvae and evaluate the "auditory response". They found a significative decrease in the response (movement trajectory, swimming velocity, distance) in the hsd17b7 mutant. The authors conclude that this gene is crucial for the "auditory function in zebrafish".

This is an overstatement:

(a) First, this test is adequate as a screening tool to identify animals that have lost completely the behavioral response to this acoustic and vibrational stimulation, which also involves a motor response. However, additional tests are required to confirm an auditory origin of the defect, such as Auditory Evoked Potential recordings, or for the vestibular function, the Vestibulo-Ocular Reflex.

(b) Secondly, the behavioral defects observed in the mutant compared to the control are significantly different, but the differences are slight, contained within the Standard Deviation (20% for velocity, 25% for distance). To this point, the Figure 2 B and C plots are misleading because their y-axis do not start at 0.

(8) Overexpression of HSD17B7 in cell line HEI-OC1 apparently "significantly increases" the intensity of cholesterol-related signal using a genetically encoded fluorescent sensor (D4H-mCherry). However, the description of this quantification (per cell or per surface area) and the normalization of the fluorescent signal are not provided.

(9) When this experiment is conducted in vivo in zebrafish, a reduction in the "DH4 relative intensity" is detected (same issue with the absence of a detailed method description). However, as the difference is smaller than the standard deviation, this raises questions about the biological relevance of this result.

(10) The authors identified a deaf child as a carrier of a nonsense mutation in HSB17B7, which is predicted to terminate the HSB17B7 protein before the transmembrane domain. However, as no genetic linkage is possible, the causality is not demonstrated.

(11) Previous results obtained from mouse HSD17B7-KO (citation below) are not described in sufficient detail. This is critical because, in this paper, the mouse loss-of-function of HSD17B7 is embryonically lethal, whereas no apparent phenotype was reported in heterozygotes, which are viable and fertile. Therefore, it seems unlikely that heterozygous mice exhibit hearing loss or vestibular defects; however, it would be essential to verify this to support the notion that the truncated allele found in one patient is causal.

Hydroxysteroid (17beta) dehydrogenase 7 activity is essential for fetal de novo cholesterol synthesis and for neuroectodermal survival and cardiovascular differentiation in early mouse embryos.

Jokela H, Rantakari P, Lamminen T, Strauss L, Ola R, Mutka AL, Gylling H, Miettinen T, Pakarinen P, Sainio K, Poutanen M.

Endocrinology. 2010 Apr;151(4):1884-92. doi: 10.1210/en.2009-0928. Epub 2010 Feb 25.

(12) The authors used this truncated protein in their startle response and FM4-64 assays. First, they show that contrary to the WT version, this truncated form cannot rescue their phenotypes when overexpressed. Secondly, they tested whether this truncated protein could recapitulate the startle reflex and FM4-64 phenotypes of the mutant allele. At the homozygous level (not mentioned by the way), it can apparently do so to a lesser degree than the previous mutant. Again, the differences are within the Standard Deviation of the averages. The authors conclude that this mutation found in humans has a "negative effect" on hearing, which is again not supported by the data.

(13) The authors looked at the distribution of the HSB17B7 in a cell line. The WT version goes to the ER, while the truncated one forms aggregates. An interesting experiment consisted of co-expressing both constructs (Figure S6) to see whether the truncated version would mislocalize the WT version, which could be a mechanism for a dominant phenotype. However, this is not the case.

(14) Through mass spectrometry of HSB17B7 proteins in the cell line, they identified a protein involved in ER retention, RER1. By biochemistry and in a cell line, they show that truncated HSB17B7 prevents the interaction with RER1, which would explain the subcellular localization.

Hydroxysteroid (17beta) dehydrogenase 7 activity is essential for fetal de novo cholesterol synthesis and for neuroectodermal survival and cardiovascular differentiation in early mouse embryos.

Jokela H, Rantakari P, Lamminen T, Strauss L, Ola R, Mutka AL, Gylling H, Miettinen T, Pakarinen P, Sainio K, Poutanen M.

Endocrinology. 2010 Apr;151(4):1884-92. doi: 10.1210/en.2009-0928. Epub 2010 Feb 25.

(15) Information and specificity validation of the HSB17B7 antibody are not presented. It seems that it is the same used on mice by IF and on zebrafish by Western. If so, the antibody could be used on zebrafish by IF to localize the endogenous protein (not overexpression as done here). Secondly, the specificity of the antibody should be verified on the mutant allele. That would bring confidence that the staining on the mouse is likely specific.

Author response:

The following is the authors’ response to the original reviews.

Public Reviews:

Reviewer #1 (Public review):

(1) The pathogenic mechanism of the E182STOP variant is unclear. The mutant protein does not appear to affect WT protein localization, arguing against a dominant-negative effect. Yet, overexpression of HSD17B7-E182* alone causes toxicity in zebrafish and mislocalizes cholesterol in HEI-OC1 cells, suggesting a gain-of-function or toxic effect. In addition, the variant mRNA is expressed at a low level, consistent with nonsense-mediated decay. This apparent complexity and inconsistency need clearer explanation.

We appreciate the reviewer’s careful evaluation of this mechanistic complexity. Based on our combined molecular, cellular, and in vivo data, we propose that the pathogenic effect of the HSD17B7-E182* variant reflects a composite mechanism, rather than a classical dominant-negative effect.

At the transcript level, the E182* variant introduces a premature termination codon and shows markedly reduced mRNA abundance, consistent with partial degradation by nonsense-mediated mRNA decay. This reduction is expected to decrease overall HSD17B7 dosage, contributing a loss-of-function component. Unlike HSD17B7, the truncated HSD17B7E182* mislocalizes cholesterol in HEI-OC1 cells, and overexpression alone reduces hair cell MET function and startle response in zebrafish embryos. We therefore propose that the truncated protein disturbing local cholesterol homeostasis, thereby exerts a toxic or ectopic gain-of-function.

We have revised the manuscript to clarify the dual-mechanism model.

(2) The link to human deafness is based on a single heterozygous patient with no syndromic features. Given that nearly all known cholesterol metabolism disorders are syndromic, this raises concerns about causality or specificity. The term "novel deafness gene" is premature without additional cases or segregation data.

We thank the reviewer for this important point. We fully agree that, based on a single heterozygous case without segregation data, it is premature to designate HSD17B7 as a novel deafness gene. Therefore, we have revised the manuscript to use the description of "candidate deafness genes".

(3) The localization of HSD17B7 should be clarified better: In HEI-OC1 cells, HSD17B7 localizes to the ER, as expected. In mouse hair cells, the staining pattern is cytosolic and almost perfectly overlaps with the hair cell marker used, Myo7a. This needs to be discussed. Without KO tissue, HSD17B7 antibody specificity remains uncertain.

We thank the reviewer for the constructive comments regarding HSD17B7 localization and antibody specificity.

Regarding subcellular localization, the original Figure 1K was intended to demonstrate the expression of HSD17B7 in mouse hair cells. To address this concern, we performed additional immunostaining on dissected organ of Corti sections at P1, P4, and P7 using higher magnification. Using parvalbumin as a hair cell marker, HSD17B7 displayed a partially punctate intracellular pattern in hair cells (revised Figure 1K). This pattern is consistent with localization to membrane-associated compartments, including the endoplasmic reticulum, and agrees with the ER-associated localization observed in HEI-OC1 cells and zebrafish hair cells. In mature hair cells, ER-associated signals may appear cytosolic and overlap with general hair cell markers such as Myo7a.

Regarding antibody specificity, although HSD17B7 knockout tissue was not available, we performed complementary validation experiments in HEI-OC1 cells. Cells were transfected with pCMV-Flag, pCMV-Flag-hHSD17B7WT, or pCMV-hHSD17B7WT-EGFP constructs and stained with anti-Flag, anti-EGFP, and anti-HSD17B7 antibodies. The HSD17B7 antibody signal showed strong co-localization with both FLAG- and EGFP-tagged HSD17B7 (revised Figure S1A and B), supporting its specificity.

Reviewer #2 (Public review):

(1) The statement that HSD17B7 is "highly" expressed in sensory hair cells in mice and zebrafish seems incorrect for zebrafish:

(a) The data do not support the notion that HSB17B7 is "highly expressed" in zebrafish. Compared to other genes (TMC1, TMIE, and others), the HSB17B7 level of expression in neuromast hair cells is low (Figure 1F), and by extension (Figure 1C), also in all hair cells. This interpretation is in line with the weak detection of an mRNA signal by ISH (Figure 1G I"). On this note, the staining reported in I" does not seem to label the cytoplasm of neuromast hair cells. An antisense probe control, along with a positive control (such as TMC1 or another), is necessary to interpret the ISH signal in the neuromast.

We thank the reviewer for this detailed evaluation and agree that the description of HSD17B7 expression in zebrafish hair cells requires clarification.

To address this, we performed a quantitative comparison of average expression levels within neuromast hair cells using log-normalized single-cell RNA-seq data. This analysis shows that hsd17b7 is expressed at a level comparable to several known MET-associated genes (e.g., tmc1 and lhfpl5a) (revised Figure 1D). Regarding the pseudotime heatmap (Figure 1F), we now state that this analysis illustrates temporal expression dynamics within neuromast hair cell development.

In addition, we have clarified the interpretation of the whole-mount in situ hybridization data by emphasizing that the signal indicates spatial enrichment rather than high transcript abundance.

We have updated the figure panels, legends, and corresponding text in the Results section to reflect these changes.

(b) However, this is correct for mouse cochlear hair cells, based on single-cell RNA-seq published databases and immunostaining performed in the study. However, the specificity of the anti-HSD17B7 antibody used in the study (in immunostaining and western blot) is not demonstrated. Additionally, it stains some supporting cells or nerve terminals. Was that expression expected?

To assess antibody specificity, we performed validation experiments using distinct epitopes. In HEI-OC1 cells transfected with pCMV-Flag-HSD17B7, or pCMV-HSD17B7-EGFP constructs, immunostaining with anti-HSD17B7 showed strong co-localization with both FLAG- and EGFP-tag (revised Figure S1B). In addition, western blot analyses using the same constructs confirmed the specific detection of HSD17B7 protein (revised Figure S1B). These validation data have now been included as supplementary figures in the revised manuscript and provide independent supporting evidence for the specificity of the anti-HSD17B7 antibody.

(2) A previous report showed that HSD17B7 is expressed in mouse vestibular hair cells by single-cell RNAseq and immunostaining in mice, but it is not cited: Spatiotemporal dynamics of inner ear sensory and non-sensory cells revealed by single-cell transcriptomics. Jan TA, Eltawil Y, Ling AH, Chen L, Ellwanger DC, Heller S, Cheng AG. Cell Rep. 2021 Jul 13;36(2):109358. doi: 10.1016/j.celrep.2021.109358.

We have now cited this reference in the revised manuscript.

(3) Overexpressed HSD17B7-EGFP C-terminal fusion in zebrafish hair cells shows a punctiform signal in the soma but apparently does not stain the hair bundles. One limitation is the consequence of the C-terminal EGFP fusion to HSD17B7 on its function, which is not discussed.

We thank the reviewer for raising this important technical point. The apparent absence of an HSD17B7-EGFP signal in hair bundles is primarily due to the imaging strategy and the selection of representative images. In zebrafish hair cells, the EGFP signal within hair bundles is extremely strong. To better visualize the intracellular distribution of HSD17B7 within the hair cell soma, we selected representative confocal optical sections that were focused on the cell body rather than on the apical hair bundle plane. As a result, the hair bundle signal is not visible in the images shown.

Importantly, we agree that C-terminal EGFP fusion may potentially influence protein localization or function. We have therefore revised the Discussion to discuss this limitation and to clarify that our central conclusions regarding HSD17B7 function are primarily supported by loss-of-function analyses, rescue experiments using untagged mRNA, and cholesterol perturbation phenotypes, rather than relying solely on EGFP-tagged overexpression constructs.

(4) A mutant Zebrafish CRISPR was generated, leading to a truncation after the first 96 aa out of the 340 aa total. It is unclear why the gene editing was not done closer to the ATG. This allele may conserve some function, which is not discussed.

Targeting regions close to the ATG is indeed a commonly used strategy for CRISPR-mediated gene disruption. In this study, sgRNA selection was guided by online CRISPR design tools (CRISPRscan), prioritizing predicted cutting efficiency and specificity. This strategy resulted in a frameshift mutation introducing a premature stop codon after amino acid 96 of the 340-aa Hsd17b7 protein.

Importantly, this truncation removes most of the conserved catalytic core required for 17β-hydroxysteroid dehydrogenase activity, including key motifs involved in NAD(P)-binding and substrate recognition. Therefore, although the mutation does not occur immediately adjacent to the ATG, the resulting allele is predicted to lack enzymatic function. We have clarified this rationale and discussed the functional consequences of the truncation in the revised manuscript.

(5) The hsd17b7 mutant allele has a slightly reduced number of genetically labeled hair cells (quantified as a 16% reduction, estimated at 1-2 HC of the 9 HC present per neuromast). On a note, it is unclear what criteria were used to select HC in the picture. Some Brn3C:mGFP positive cells are apparently not included in the quantifications (Figure 2F, Figure 5A).

Upon re-evaluation, we recognized that the original figure annotations were not sufficiently clear and may have led to confusion regarding hair cell selection. In the original images, the absence of dashed outlines around some Brn3c:mGFP+ cells may have been misinterpreted as their exclusion from analysis. To address this issue, we have revised Figures 2F and 5A by updating the annotations to ensure that all Brn3c:mGFP+ hair cells within each neuromast are clearly visible and unambiguously included (revised Figures 2F and 6A). Corresponding figure legends have also been revised to clarify the criteria used for hair cell identification and quantification.

(6) The authors used FM4-64 staining to evaluate the hair cell mechanotransduction activity indirectly. They found a 40% reduction in labeling intensity in the HCs of the lateral line neuromast. Because the reduction of hair cell number (16%) is inferior to the reduction of FM4-64 staining, the authors argue that it indicates that the defect is primarily affecting the mechanotransduction function rather than the number of HCs. This argument is insufficient. Indeed, a scenario could be that some HC cells died and have been eliminated, while others are also engaged in this path and no longer perform the MET function. The numbers would then match. If single-cell staining can be resolved, one could determine the FM4-64 intensity per cell. It would also be informative to evaluate the potential occurrence of cell death in this mutant. On another note, the current quantification of the FM4-64 fluorescence intensity and its normalization are not described in the methods. More importantly, an independent and more direct experimental assay is needed to confirm this point. For example, using a GCaMP6-T2A-RFP allele for Ca2+ imaging and signal normalization.

We have revised the FM4-64 quantification strategy. Instead of measuring fluorescence intensity at the neuromast level, FM4-64 uptake was re-quantified at the single hair cell level. Hair cells within each neuromast were identified based on mGFP labeling, and the mean FM4-64 fluorescence intensity was measured for each individual hair cell. The average FM4-64 intensity per hair cell was then calculated for each neuromast and used for group comparisons (revised Figures 2F, 6B, and 8B, Figure S5B). The updated quantification method, normalization procedure, and analysis pipeline have now been described in the revised Methods section.

As supportive evidence, we further analyzed single-cell RNA-seq data from control and hsd17b7 mutant hair cells (revised Figure 3). This analysis revealed dysregulation of multiple genes involved in the MET machinery, including reduced expression of tip-link–associated components and altered expression of other MET-related genes. While these transcriptional changes do not constitute a direct functional assay, they are consistent with perturbation of MET-associated pathways and complement the FM4-64 findings.

(7) The authors used an acoustic startle response to elicit a behavioral response from the larvae and evaluate the "auditory response". They found a significative decrease in the response (movement trajectory, swimming velocity, distance) in the hsd17b7 mutant. The authors conclude that this gene is crucial for the "auditory function in zebrafish".

This is an overstatement:

(a) First, this test is adequate as a screening tool to identify animals that have lost completely the behavioral response to this acoustic and vibrational stimulation, which also involves a motor response. However, additional tests are required to confirm an auditory origin of the defect, such as Auditory Evoked Potential recordings, or for the vestibular function, the Vestibulo-Ocular Reflex.

We thank the reviewer for highlighting the limitations in interpreting the acoustic startle assay. We have revised the manuscript to avoid overstatement and now describe the observed phenotype as a reduction in the behavioral response to acoustic and vibrational stimulation, rather than concluding a specific impairment of auditory function.

(b) Secondly, the behavioral defects observed in the mutant compared to the control are significantly different, but the differences are slight, contained within the Standard Deviation (20% for velocity, 25% for distance). To this point, the Figure 2 B and C plots are misleading because their y-axis do not start at 0.

We have corrected Figures 2B and 2C so that the y-axes start at zero, thereby providing a more transparent visualization of the behavioral differences. The figure legends have also been revised to clarify the presentation of the data.

(8) Overexpression of HSD17B7 in cell line HEI-OC1 apparently "significantly increases" the intensity of cholesterol-related signal using a genetically encoded fluorescent sensor (D4H-mCherry). However, the description of this quantification (per cell or per surface area) and the normalization of the fluorescent signal are not provided.

The quantification of the D4H-mCherry signal in HEI-OC1 cells was performed at the single-cell level. Specifically, individual cells were segmented based on morphology, and the mean fluorescence intensity of D4H-mCherry per cell was measured. To account for variability in cell size and imaging conditions, fluorescence intensity was normalized to the background signal measured from cell-free regions in the same field of view. We have now clarified the quantification strategy and normalization procedure in the revised Methods and Results sections.

(9) When this experiment is conducted in vivo in zebrafish, a reduction in the "DH4 relative intensity" is detected (same issue with the absence of a detailed method description). However, as the difference is smaller than the standard deviation, this raises questions about the biological relevance of this result.

We have now clarified the quantification strategy and normalization procedure in the revised Methods and Results sections.

(10) The authors identified a deaf child as a carrier of a nonsense mutation in HSB17B7, which is predicted to terminate the HSB17B7 protein before the transmembrane domain. However, as no genetic linkage is possible, the causality is not demonstrated.

We thank the reviewer for raising this important point. Unfortunately, we were unable to obtain the parents' genetic testing data to perform formal genetic and linkage analysis. To address this limitation, we have revised the manuscript to avoid causal overstatement and now describe the HSD17B7 E182* variant as a candidate pathogenic variant associated with hearing loss. Importantly, our functional analyses in zebrafish and cell-based systems demonstrate that the E182* truncation abolishes key biological activities of HSD17B7, including subcellular localization, cholesterol regulation, mechanotransduction-related activity, and behavioral responses. These convergent functional data provide biological support for the potential pathogenic relevance of this variant.

(11) Previous results obtained from mouse HSD17B7-KO (citation below) are not described in sufficient detail. This is critical because, in this paper, the mouse loss-of-function of HSD17B7 is embryonically lethal, whereas no apparent phenotype was reported in heterozygotes, which are viable and fertile. Therefore, it seems unlikely that heterozygous mice exhibit hearing loss or vestibular defects; however, it would be essential to verify this to support the notion that the truncated allele found in one patient is causal.

Hydroxysteroid (17beta) dehydrogenase 7 activity is essential for fetal de novo cholesterol synthesis and for neuroectodermal survival and cardiovascular differentiation in early mouse embryos.

Jokela H, Rantakari P, Lamminen T, Strauss L, Ola R, Mutka AL, Gylling H, Miettinen T,

Pakarinen P, Sainio K, Poutanen M. Endocrinology. 2010 Apr;151(4):1884-92. doi: 10.1210/en.2009-0928. Epub 2010 Feb 25.

We thank the reviewer for raising this important point. We acknowledge that previous work has shown that complete loss of Hsd17b7 in mice is embryonically lethal, whereas heterozygous animals are viable and fertile (Jokela et al., 2010). Notably, this study primarily focused on embryonic development, cholesterol metabolism, and cardiovascular and neuroectodermal survival, and auditory or vestibular functions were not specifically examined. Therefore, subtle or sensory organ–specific phenotypes in heterozygous mice cannot be excluded.

The human variant identified in this study (E182*) is a nonsense mutation predicted to truncate the HSD17B7 protein prior to the transmembrane and cytoplasmic domains. We therefore present it as a candidate loss-of-function variant, providing supportive human genetic evidence that is consistent with our functional analyses in zebrafish hair cells, rather than as definitive proof of causality. We have revised the manuscript to clarify these points and to acknowledge this limitation.

(12) The authors used this truncated protein in their startle response and FM4-64 assays. First, they show that contrary to the WT version, this truncated form cannot rescue their phenotypes when overexpressed. Secondly, they tested whether this truncated protein could recapitulate the startle reflex and FM4-64 phenotypes of the mutant allele. At the homozygous level (not mentioned by the way), it can apparently do so to a lesser degree than the previous mutant. Again, the differences are within the Standard Deviation of the averages. The authors conclude that this mutation found in humans has a "negative effect" on hearing, which is again not supported by the data.

We thank the reviewer for this important comment. We agree that the overexpression strategy employed in this study does not fully replicate the endogenous heterozygous state observed in patients, and that the magnitude of the observed effects varies across samples. Accordingly, our experiments were not intended to demonstrate a definitive causal role of the HSD17B7 E182* variant in hearing loss.

Instead, the overexpression assays were designed to assess whether the truncated HSD17B7 protein displays abnormal cellular properties and whether its presence can interfere with processes relevant to hair cell function. Under these conditions, HSD17B7E182* exhibited aberrant subcellular localization, altered intracellular cholesterol distribution, and was associated with reduced FM4-64 uptake and changes in startle-associated behaviors, whereas the wild-type protein did not.

We revised the manuscript to moderate our conclusions. Rather than claim that the E182* mutation has a definitive “negative effect on auditory function,” we now describe it as a functionally compromised allele that disrupts cholesterol distribution and MET-related activity under overexpression conditions, providing mechanistic support consistent with our zebrafish loss-of-function data and the identification of this variant in a patient with hearing loss. In addition, the "negative effect" statement was based on the result that overexpression of the E182* mutation in wild-type embryos caused the compromised MET function and startle response defect.

(13) The authors looked at the distribution of the HSB17B7 in a cell line. The WT version goes to the ER, while the truncated one forms aggregates. An interesting experiment consisted of co-expressing both constructs (Figure S6) to see whether the truncated version would mislocalize the WT version, which could be a mechanism for a dominant phenotype. However, this is not the case.

We thank the reviewer for raising this important point regarding a potential dominant-negative mechanism. Consistent with the reviewer’s interpretation, we found that HSD17B7WT predominantly localizes to the endoplasmic reticulum, whereas the truncated HSD17B7E182* protein forms intracellular aggregates. Importantly, we further observed that the E182* mutation markedly reduces the stability of both HSD17B7 mRNA and protein, resulting in substantially decreased abundance of the truncated protein (Figure S6B–E). As a consequence, the cellular levels of HSD17B7^E182* are abnormally low.

Based on these findings, we consider it unlikely that the E182* variant exerts its effect through interference with the wild-type protein. Our results suggest that the heterozygous c.544G>T (p.E182*) variant contributes to auditory dysfunction through potential pathogenic mechanisms: 1, haploinsufficiency caused by reduced HSD17B7 expression, 2, functional impairment due to altered protein subcellular localization and cholesterol distribution.

We have revised the Results and Discussion sections. Our conclusions now emphasize that the functional impact of this variant is attributable to decreased effective HSD17B7 dosage, consistent with the observed defects in cholesterol synthesis, MET-related activity, and auditory-associated phenotypes in our model.

(14) Through mass spectrometry of HSB17B7 proteins in the cell line, they identified a protein involved in ER retention, RER1. By biochemistry and in a cell line, they show that truncated HSB17B7 prevents the interaction with RER1, which would explain the subcellular localization.

Consistent with the reviewer’s interpretation, wild-type HSD17B7 interacts with RER1, a protein known to participate in ER retention, whereas this interaction is lost in the truncated HSD17B7 variant. We propose that RER1 is an interacting partner of HSD17B7, providing a mechanistic explanation for the protein's subcellular localization.

(15) Information and specificity validation of the HSB17B7 antibody are not presented. It seems that it is the same used on mice by IF and on zebrafish by Western. If so, the antibody could be used on zebrafish by IF to localize the endogenous protein (not overexpression as done here). Secondly, the specificity of the antibody should be verified on the mutant allele. That would bring confidence that the staining on the mouse is likely specific.

We thank the reviewer for raising this important point regarding antibody specificity and validation. Information on the HSD17B7 antibody and its validation has been provided in our response to comment 1, where we described the use of antibodies recognizing different epitopes and the experimental strategies employed to assess specificity (revised Figure S1A and B).

Although the same antibody was used for Western blot analysis in zebrafish samples, its performance in immunofluorescence staining of zebrafish tissues was suboptimal, with relatively high background. For this reason, we did not rely on this antibody for endogenous Hsd17b7 localization in zebrafish by immunofluorescence and instead employed tagged constructs for subcellular localization analyses. This approach provides more reliable and interpretable localization information under the current experimental conditions.

Recommendations for the authors:

Reviewing Editor Comments:

Suggested revisions to help improve the study and the eLife Assessment:

(1) FM4-64 uptake: Isolate the effect of hair cell loss and MET reduction.

(2) Clarify the mechanistic model: Is the mutant protein pathogenic due to toxicity, lack of expression or function, or both? Come up with a clearer causal chain of events.

(3) Mouse immunostaining: Validate the HSD17B7 antibody, and since mouse RNAseq data (gEAR database) suggest that HSD17B7 expression increases dramatically between P0-P5, show this developmental progression by immunostaining of the mouse organ of Corti at P0, P3, and P5.

(4) The HSD17B7-E182* expression disrupts cholesterol (D4H staining) in OC1 cells. This should also be demonstrated in the mutant zebrafish.

(5) Structural modeling of E182* is uninformative; half the protein is absent. This kind of analysis is better suited for missense variants. Suggest removing this analysis.

We thank the Reviewing Editor for these constructive suggestions. The major points raised here substantially overlap with the concerns raised in the public reviews. In response, we have:

(1) revised FM4-64 quantification and interpretation to better distinguish hair cell loss from MET impairment;

(2) Clarify the mechanistic mode. Mechanistically, the mutation decreases mRNA abundance and significantly reduces protein levels. Moreover, expression of the p.E182* mutation disrupted the interaction between HSD17B7 and the ER retention receptor RER1, leading to aberrant subcellular localization and altered cholesterol distribution, thereby exacerbating HC dysfunction.

(3) provided additional validation of the HSD17B7 antibody using antibodies targeting distinct epitopes, and extended mouse organ of Corti immunostaining to postnatal stages P1, P4, and P7 to demonstrate the developmental upregulation of HSD17B7 expression;

(4) added in vivo zebrafish experiments demonstrating that expression of HSD17B7E182* disrupts cholesterol distribution in hair cells, consistent with the effects observed in HEI-OC1 cells using D4H staining;

(5) removed the structural modeling of the E182* variant.

Recommendations for the authors:

The recommendations from Reviewer #1 and Reviewer #2 were carefully considered and addressed. Most of these points overlap with the public reviews and the Reviewing Editor's comments and have been addressed through a revised mechanistic interpretation, additional clarifications in the Methods, more moderate claims regarding auditory function and human genetics, and the removal or revision of potentially misleading analyses. In addition, a number of minor issues were corrected, including missing or incorrect references, repetitive or unclear statements in the Introduction, insufficient methodological details, imprecise terminology, and typographical or formatting errors. Collectively, these revisions improve the clarity, rigor, and transparency of the study without altering its central conclusions.

Version 2

Abstract

Cholesterol homeostasis is fundamental to cellular function, and its disruption underlies a wide range of human diseases. However, the contribution of cholesterol biosynthesis to auditory physiology remains poorly understood. HSD17B7 (17β-Hydroxysteroid dehydrogenase type 7) catalyzes the conversion of zymosterone to zymosterol, a key step in the post-lanosterol cholesterol biosynthetic pathway. Here, we found that Hsd17b7 is highly enriched in sensory hair cells of zebrafish and mice. The deficiency of Hsd17b7 deficiency reduced intracellular cholesterol levels in HEI-OC1 cells and zebrafish hair cells, thereby compromising MET and acoustic startle responses. A heterozygous nonsense variant (c.544G>T; p.E182*) in HSD17B7 was identified in an individual with bilateral profound hearing loss. mRNA of c.544G>T HSD17B7 failed to rescue the impaired MET and acoustic startle response of hsd17b7 mutants. Mechanistically, the mutation decreases mRNA abundance and significantly reduces protein. Moreover, expression of the p.E182* mutation disrupted the interaction between HSD17B7 and the ER retention receptor RER1, leading to aberrant subcellular localization and altered cholesterol distribution, thereby exacerbating HC dysfunction. Together, our findings suggest a conserved and essential role for HSD17B7-mediated cholesterol biosynthesis in sensory hair cell function and identify HSD17B7 as a candidate gene for sensorineural hearing loss.

Introduction

Hearing loss is one of the most common sensory disorders worldwide and represents a major global health burden. According to the World Health Organization, more than 466 million people currently suffer from disabling hearing loss, and this number is expected to increase substantially in the coming decades [1]. Genetic factors account for approximately 60% of congenital hearing loss, and mutations in genes that regulate the development and function of sensory hair cells (HCs) are the main cause of hereditary deafness. Although more than 150 non-syndromic deafness genes have been identified [2], the molecular mechanisms underlying HC dysfunction remain incompletely understood, and many pathogenic genes remain to be discovered.

HCs are the primary mechanoreceptors of the auditory and vestibular systems, converting mechanical stimuli into electrical signals through mechanotransduction (MET). This process critically depends on the integrity and biophysical properties of hair cell membranes, particularly those of the stereocilia. Cholesterol is the most abundant sterol molecule in mammalian cells and plays essential roles in maintaining membrane structure, synthesizing important hormones, facilitating synapse formation, and mediating cell signaling transduction [3–6]. Previous studies have shown that dysregulated intracellular cholesterol homeostasis contributes to auditory defects [6–8], including hereditary [9–13], noise-induced [14–16], ototoxic [17, 18], and age-related hearing loss [19–21]. HSD17B7 is a member of the hydroxysteroid dehydrogenase family and functions as a key enzyme in the cholesterol biosynthetic pathway by catalyzing the conversion of zymosterone to zymosterol. In addition to its role in steroid metabolism [22] and its involvement in hormone-related cancers [23, 24], HSD17B7 has been reported to be expressed in multiple tissues, including the brain, eye, and inner ear [25, 26]. Notably, single-cell RNA sequencing (scRNA-seq) and immunostaining analyses have suggested that HSD17B7 is enriched in mouse vestibular HCs [27]. Despite this expression pattern, the functional role of HSD17B7 in HCs and MET has not been investigated, and no pathogenic variants have been reported to be associated with hearing loss.

Here, our findings reveal a previously unrecognized link between hair cell–intrinsic cholesterol biosynthesis and the function of sensory hair cells, and identify HSD17B7 as a candidate gene underlying sensory hearing disorders.

Results

Hsd17b7 is expressed in sensory hair cells of zebrafish and mice

To assess the role of Hsd17b7 in the auditory system, we first investigated the expression of hsd17b7 in the developing zebrafish. scRNA-seq data analysis (accession no. GSE221471) [28] categorized lateral line hair cells (LLHCs) (cluster 0), supporting cell (cluster 1), macula hair cells (MHCs) (cluster 2), crista hair cells (CHCs) 1 (cluster 3), mantle cell (cluster 4), and CHCs 2 (cluster 5) using the Seurat 4.0.1 platform (Figure 1A). The feature plot and violin plot showed that hsd17b7 was expressed in sensory HCs, especially in LLHCs and CHCs (Figure 1B and C). Quantitative comparison of average expression levels within LLHCs indicates that hsd17b7 is expressed at a level comparable to several known MET-associated genes (e.g., tmc1 and lhfpl5a) (Figure 1D). To examine the temporal dynamics of hsd17b7 expression during LLHCs differentiation, LLHCs, supporting cells, and mantle cells were ordered along pseudotime using Monocle 3 (Figure 1E). A heatmap was generated to visualize the expression dynamics of marker genes along the LLHC developmental trajectory (Figure 1F). These analyses suggested a gradual increase in hsd17b7 expression during LLHC maturation (Figure 1E and F). Consistently, whole-mount in situ hybridization (WISH) further showed that hsd17b7 was specifically enriched in neuromast and crista regions at 72 and 96 hours post-fertilization (hpf) (Figure 1G). To further explore the subcellular localization of hsd17b7 in zebrafish HCs, we generated a hsd17b7-egfp fusion construct driven by a HC-specific myo6b promoter [29, 30]. Confocal imaging revealed that hsd17b7-EGFP localized predominantly to the cytoplasm in both CHCs and LLHCs, exhibiting a punctate distribution pattern (Figure 1H and I).

Conserved expression of Hsd17b7 in zebrafish and mice sensory hair cells.

(A) The UMAP analysis of the zebrafish scRNA-seq data. LLHCs, lateral line hair cells; MHCs, macula hair cells; CHCs, crista hair cells; (B, C) Feature plot and violin plot of hsd17b7 expression across cell types. (D) Quantitative comparison of average expression levels of hsd17b7 within LLHCs. (E) hsd17b7 expression increased along the pseudotime trajectory of LLHC formation. (F) Heatmap of marker genes along the pseudotime trajectory. (G) The expression of the hsd17b7 in 72 hpf and 96 hpf embryos was detected by whole-mount in situ hybridization analysis. Dashed circles indicate the otic vesicle and the neuromast. (H, I) Representative images of the CHCs and LLHCs in Tg(myo6b: hsd17b7-egfp) at 4 dpf. (J) The average expression of Hsd17b7 and HC marker genes across mouse HC subtypes from published scRNA-seq datasets. UHCs, utricular hair cells. (K) Immunostaining and quantification analysis of HSD17B7 in dissected mouse organ of Corti sections at P1, P4, and P7. Dashed circles indicate outer hair cells (OHCs) and inner hair cells (IHCs).

To further elucidate Hsd17b7 expression in the mammalian auditory system, we analyzed three published scRNA-seq datasets from mouse inner ear tissues [31–33]. These analyses revealed that Hsd17b7 was enriched in sensory hair cells, including outer hair cells (OHCs), inner hair cells (IHCs), and utricular hair cells (UHCs), while lower expression levels were observed in CHCs (Figure 1J). Next, we performed immunostaining to assess Hsd17b7 protein expression in the organ of Corti. Hsd17b7 was detected in both OHCs and IHCs at P1, P4, and P7 (Figure 1K). The specificity of the Hsd17b7 antibody was further supported by independent validation experiments in HEI-OC1 cells (Figure S1A and B). These results indicate that Hsd17b7 is expressed and enriched in HCs in both zebrafish and mice.

In addition, phylogenetic and sequence alignment analysis revealed that vertebrate HSD17B7s share a significant similarity in amino acid sequences (Figure S2A). UniProt sequence alignment analysis revealed that Danio rerio hsd17b7 (NP_001070796.1) shares 58.2% identity with Homo sapiens HSD17B7 (NP_057455.1) and 67.1% identity with Mus musculus Hsd17b7 (NP_034606.3), respectively (Figure S2B). These data suggest that Hsd17b7 is conserved and expressed in HCs across vertebrate species.

Hsd17b7 deficiency impaired acoustic startle responses and MET activity in zebrafish

To investigate the hsd17b7 function in the auditory system, we generated hsd17b7 mutant alleles using CRISPR/Cas9-mediated genome editing (Figure S3A–D). Sequence analysis identified multiple mutant alleles, including a 4-bp deletion and two 6-bp deletions in exon 3 (Figure S3E). The 4-bp deletion allele was selected for further analysis, as it led to a frameshift and premature stop codon, producing a severely truncated protein lacking most of the conserved catalytic domain required for hsd17b7 enzymatic activity (Figure S3F).

We first examined acoustic startle responses [34] at 5 dpf (Figure 2A). Compared with control larvae, hsd17b7 mutants exhibited significantly reduced movement trajectories, peak swimming velocity, and total travel distance following acoustic stimulation (Figure 2A–C), indicating impaired startle response. Microinjection of hsd17b7 mRNA into mutant embryos significantly restored these behavioral phenotypes (Figure 2A–C), confirming that the phenotype was specifically caused by loss of Hsd17b7.

Loss of hsd17b7 impaired acoustic startle responses and mechanotransduction activity in zebrafish.

(A) Left, schematic of the experimental setup used to assess acoustic startle responses in zebrafish larvae. Right, representative locomotor trajectories showing C-start responses to a single acoustic stimulus (9 dB re. 1 m·s⁻², 60 Hz tone burst) in control, hsd17b7 knockout (KO), and hsd17b7 mRNA–rescued larvae at 5 dpf. Scale bars, 10 mm. (B, C) Quantification of peak swimming velocity (B) and total movement distance (C) in response to acoustic stimulation shown in (A) (n = 20). One-way ANOVA followed by Tukey’s multiple comparisons, *P < 0.05, **P < 0.01, ***P < 0.001, ns, non-significant (p > 0.05), mean ± SEM. (D) Representative confocal images of lateral line hair cells (LLHCs, green) in Tg(Brn3c:mGFP) larvae at 4 dpf from control, hsd17b7 KO, and hsd17b7 mRNA–rescued groups. Scale bars, 20 μm. (E) Quantification of HC number per neuromast (n = 30). One-way ANOVA followed by Tukey’s multiple comparisons, *P < 0.05, **P < 0.01, ns, non-significant (p > 0.05), mean ± SD. (F) Representative images of LLHCs (green) and FM4-64–labeled functional HCs (red) in single neuromasts of Tg(Brn3c:mGFP) larvae at 5 dpf from the indicated groups. White dashed outlines indicate LLHCs. Scale bars, 20 μm. (G) Quantification of relative FM4-64 fluorescence intensity per neuromast (n = 25). One-way ANOVA followed by Tukey’s multiple comparisons, ****P < 0.0001, ns, non-significant (p > 0.05), mean ± SD.

To validate these findings, antisense morpholino oligonucleotides (hsd17b7 Mo) targeting the junction of exon 1/intron 1 were designed to knock down hsd17b7 expression (Figure S4A and B). RT-PCR and immunoblotting analyses confirmed effective disruption of normal splicing and a substantial reduction in Hsd17b7 protein levels, both of which were rescued by co-injection of hsd17b7 mRNA (Figure S4C and D). Consistent with the mutant phenotype, hsd17b7 morphants displayed significantly reduced acoustic startle responses at 5 dpf, including decreased movement trajectories, swimming distance, and peak velocity, all of which were effectively rescued by hsd17b7 mRNA injection (Figure S4E–G).

In the Tg(Brn3c: mGFP) background [28, 35], confocal imaging revealed a significant reduction in the number of LLHCs in hsd17b7 mutants (∼16% reduction) (Figure 2D and E). To assess whether hsd17b7 loss affects the MET activity, we performed FM4-64 uptake assays, which label HCs through functional MET channels [26, 32, 33]. Quantification of FM4-64 fluorescence intensity per HC revealed a pronounced reduction in MET-dependent dye uptake in hsd17b7 mutant HCs at 5 dpf (∼40%; Figure 2F and G). A similar decrease was observed in hsd17b7 morphants (∼25%; Figure S5A and B). Importantly, hsd17b7 mRNA injection restored both FM4-64 uptake and HC number in mutants and morphants (Figure 2D–G, Figure S5A and B), indicating that hsd17b7 is required for normal MET activity in HCs.

Hsd17b7 deficiency disrupted cholesterol-associated transcriptional programs in hair cells

To investigate intrinsic transcriptional changes in HCs caused by hsd17b7 deficiency, we performed scRNA-seq on FACS-isolated mGFP+ HCs from control and hsd17b7 mutant (Figure 3A). Clustering analysis identified major inner ear (IE) and lateral line (LL) cell populations, including HCs and SCs (Figure 3B; Figure S6 A and B). Focusing on LLHCs, pseudobulk differential expression analysis revealed extensive transcriptional remodeling in hsd17b7 mutants, with 1355 genes differentially expressed, including 853 upregulated (63%) and 502 downregulated (37%) genes relative to controls (Figure 3C). Gene Ontology enrichment analysis of differentially expressed genes highlighted pathways related to lipid metabolism, membrane organization, and sensory perception, indicating broad alterations in cellular homeostasis upon disruption of cholesterol biosynthesis (Figure 3D and E).

Hsd17b7 deficiency disrupted cholesterol-associated transcriptional states in hair cells.

(A) Schematic overview of the scRNA-seq workflow. mGFP+ hair cells were isolated from control and hsd17b7 mutant zebrafish by fluorescence-activated cell sorting (FACS) and subjected to 10× Genomics single-cell RNA sequencing. (B) UMAP visualization of scRNA-seq data from control and hsd17b7 mutant samples, colored by annotated cell types, including inner ear (IE) and lateral line (LL) hair cells (HCs) and supporting cells (SCs). (C) Volcano plot showing pseudobulk differential gene expression analysis of LLHCs from hsd17b7 mutants relative to controls. Genes with adjusted p < 0.05 are highlighted, with upregulated genes shown in red and downregulated genes shown in blue. (D and E) Gene Ontology (GO) biological process enrichment analysis of genes upregulated and downregulated in hsd17b7 mutant HCs. A Sankey diagram illustrates representative genes contributing to each enriched GO term. (F) Heatmap and module score analysis of gene sets related to tip-link structure, mechanotransduction (MET), and cholesterol metabolism in LLHCs and IEHCs. Module scores were calculated using predefined gene sets, and statistical significance was assessed using the Wilcoxon rank-sum test. (G) Ridge plots showing the distribution of z-scored module scores associated with tip link structure, cholesterol uptake, and cholesterol efflux in control and hsd17b7 mutant LLHCs and IEHCs, revealing population-wide transcriptional shifts rather than discrete subpopulation changes. (H) Gene set enrichment analysis (GSEA) demonstrating significant downregulation of cholesterol biosynthesis–associated pathways in hsd17b7 mutant LLHCs.

To further assess cholesterol- and MET-associated transcriptional programs, we performed module score analysis in both LLHCs and IEHCs. In mutant IEHCs, MET, cholesterol biosynthesis, and cholesterol uptake-related gene sets were significantly upregulated, whereas gene modules associated with tip-link components and cholesterol efflux were concurrently reduced (Figure 3F and G).

In mutant LLHCs, module score analysis revealed a partially overlapping transcriptional response. Gene modules associated with tip-link components and cholesterol efflux were reduced, whereas modules representing the core MET machinery and cholesterol biosynthesis enzymes did not show significant global changes, consistent with selective remodeling rather than uniform pathway activation or repression. At the gene level, cholesterol regulatory programs appeared partially uncoupled, with downregulation of sterol-sensing transcriptional regulators (e.g., srebf2, insig1, and mbtps1) accompanied by variable expression changes among cholesterol biosynthetic enzymes. In line with this pattern, gene set enrichment analysis (GSEA) indicated an overall attenuation of cholesterol biosynthesis–associated pathways in mutant LLHCs (Figure 3H).

Together, these findings indicate that hsd17b7 deficiency induces selective and context-dependent remodeling of cholesterol- and MET-associated transcriptional programs across HC populations.

Hsd17b7 is required for maintaining cholesterol levels in hair cells

Hsd17b7 converts zymosterone to zymosterol, participating in cholesterol biosynthesis (Figure 4A) [26]. Given that both elevated and reduced cholesterol levels are detrimental to auditory function [7, 36], we first verified whether hsd17b7 regulates cholesterol homeostasis in hair cells.

Hsd17b7 deficiency reduced cholesterol in hair cells.

(A) Schematic illustration of the cholesterol biosynthesis pathway highlighting the role of HSD17B7. (B) Western blot analysis of Hsd17b7 protein levels in control and Hsd17b7-knockdown HEI-OC1 cells; quantification is shown on the right (n = 6). Significance was determined by an unpaired two-tailed Student’s t-test. ****P < 0.0001, mean ± SD. (C) Quantification of the relative cellular cholesterol levels in control and Hsd17b7-knockdown HEI-OC1 cells (n=12). Significance was determined by an unpaired two-tailed Student’s t-test. ***P < 0.001, mean ± SD. (D) Representative immunofluorescence images showing the cholesterol probe D4H-mCherry in HEI-OC1 cells transfected with pCMV-Flag or pCMV-Flag–HSD17B7 plasmids. White dashed outlines indicate transfected cells. Scale bars, 20 μm. (E) Quantification of relative D4H-mCherry fluorescence intensity for (D) (n = 16). Significance was determined by an unpaired two-tailed Student’s t-test. ****P < 0.0001, mean ± SD. (F) Experimental schematic illustrating expression of D4H-mCherry in control and hsd17b7 knockout zebrafish larvae using the Tg(Brn3c:mGFP) background. (G) Representative confocal images of D4H-mCherry (red) in crista hair cells (CHCs) and lateral line hair cells (LLHCs) of control and hsd17b7 knockout larvae at 4 dpf. White dashed outlines indicate hair cells. Scale bars, 20 μm. (H) Quantification of relative D4H-mCherry fluorescence intensity in vivo for (G) (n = 30). Significance was determined by an unpaired two-tailed Student’s t-test. ***P < 0.001, ****P < 0.0001, mean ± SD.

In vitro, knockdown of Hsd17b7 in HEI-OC1 cells resulted in a significant reduction in total cellular cholesterol levels (Figure 4B and C). To further examine cholesterol distribution, we employed the D4H cholesterol-binding probe derived from perfringolysin O, which selectively labels cholesterol at the cytoplasmic leaflet of the plasma membrane [37, 38]. Overexpression of human HSD17B7 markedly increased D4H-mCherry fluorescence intensity in HEI-OC1 cells compared with controls (Figure 4D and E), indicating that HSD17B7 positively regulates cellular cholesterol abundance.

To further assess whether hsd17b7 regulates cholesterol in vivo, we expressed D4H-mCherry in control and hsd17b7 mutant zebrafish hair cells using the Tg(Brn3c:mGFP) background (Figure 4F). Consistent with previous reports [37], D4H-mCherry was enriched in the hair bundle (Figure 4G). Notably, cholesterol levels, as indicated by D4H-mCherry intensity, were significantly reduced in both CHCs and LLHCs of hsd17b7 mutants at 4 dpf (Figure 4G and H). A comparable reduction was observed in hsd17b7 morphants (Figure S5C–F).

Together, these results demonstrate that hsd17b7 is required for maintaining cholesterol levels in HCs, both in vitro and in vivo. Given the essential role of cholesterol homeostasis in MET and auditory function [7,36,39], loss of Hsd17b7 likely compromises MET activity and acoustic startle responses by disrupting cholesterol availability in HCs.

Identification of a candidate HSD17B7 nonsense variant associated with hearing loss

Given the established role of Hsd17b7 in HCs, we investigated whether variants in the human HSD17B7 might be associated with hearing loss. Whole-genome sequencing (WGS) was performed in individuals with hearing loss who remained undiagnosed after screening of 201 known deafness-associated genes [39]. We identified a heterozygous nonsense variant in HSD17B7 (NM_016371.4), c.544G>T (p.E182*), in a child with bilateral profound congenital hearing loss (Figure 5A–C). The proband failed the newborn hearing screening, required special education during early childhood, and was diagnosed with bilateral profound hearing loss by 8 years of age. Physical examination revealed two preauricular appendages anterior to the right ear, without additional systemic abnormalities. The proband Ⅱ 1’s parents had normal hearing and no family history of hearing loss, but declined genetic testing, precluding segregation analysis (Figure 5A and B).

Identification of a heterozygous nonsense variant in human HSD17B7.

(A) Two-generation family pedigree for the affected individual. The hearing-impaired individual is indicated by a black circle (female). (B) Pure-tone audiometry audiograms for the hearing-impaired individual at 8 years old. Blue represents the results for the left ear and red for the right ear. The affected individual shows severe-to-profound or profound HI. (C) Sequence of HSD17B7 mRNA in wild type and heterozygote. (D) Multiple sequence alignment of 8 different species using the TBtools program. The p.Glu182 residue, as indicated by a red arrow, is evolutionarily conserved from zebrafish to human. (E) The domain structure of human HSD17B7 and HSD17B7E182*. ED, extracellular domain; TM, transmembrane; CD, cytoplasmic domain. The residue numbers are labeled at right.

The p.E182* variant introduces a premature stop codon at position Glu182, resulting in a truncated HSD17B7 protein (Figure 5C). The multiple sequence alignment showed that Glu182 is evolutionarily conserved across vertebrates, from zebrafish to humans (Figure 5D). According to UniProt annotation, HSD17B7 contains an extracellular domain (1-229 aa), a transmembrane region (230-250 aa), and a cytoplasmic domain (251-341 aa). The p.E182* variant is predicted to remove the transmembrane region and cytoplasmic domain (Figure 5E). Together, these analyses identify p.E182* as a candidate variant in HSD17B7 that may be associated with hearing loss.

The HSD17B7E182* truncation failed to rescue MET activity and auditory-related behaviors in hsd17b7 mutants

To assess the functional impact of the p.E182* variant, human HSD17B7 or truncation (HSD17B7E182*) mRNAs were microinjected into the fertilized one-cell stage embryos of hsd17b7 mutants. Injection of HSD17B7 mRNA significantly increased the FM4-64 uptake in LLHCs of hsd17b7 mutants, whereas the injection of HSD17B7E182* mRNA failed to restore FM4-64 labeling (Figure 6A and B), indicating that the truncated protein is unable to rescue normal MET activity. Consistently, acoustic startle response assays revealed that HSD17B7 mRNA rescued behavioral responses of hsd17b7 mutants, as assessed by movement trajectory, swimming distance, and peak velocity. In contrast, the p.E182* variant showed no rescuing effect on these behavioral deficits (Figure 6C–E).

Wild-type but not HSD17B7E182* restored MET activity and startle responses in hsd17b7 mutants.

(A) Representative images of lateral line hair cells (LLHCs, green) and FM4-64–labeled functional HCs (red) in single neuromasts of Tg(Brn3c:mGFP) at 5 dpf from control, hsd17b7 knockout (KO), KO injected with HSD17B7E182* (p.E182*) mRNA, and KO injected with HSD17B7 (WT) mRNA groups. White dashed outlines indicate LLHCs. (B) Quantification of the relative FM4-64 intensity per HC (n=25). One-way ANOVA followed by Tukey’s multiple comparisons, **P < 0.01, ****P < 0.0001, ns, non-significant (p > 0.05), mean ± SD. (C) Representative locomotor trajectories of 5 dpf larvae exhibiting behavioral responses to a single acoustic stimulus (9 dB re. 1 m·s⁻², 60 Hz tone burst) in the indicated groups. (D, E) Quantification of total movement distance (D) and peak swimming velocity (E) in response to acoustic stimulation (n = 20). One-way ANOVA followed by Tukey’s multiple comparisons, ****P < 0.0001, ns, non-significant (p > 0.05), mean ± SEM.

The HSD17B7E182* truncation disrupts intracellular cholesterol distribution and MET activity

To explore the potential contribution of HSD17B7E182* truncation to the patient’s hearing loss, we compared the cellular activation of the wild-type and truncated HSD17B7 in vitro and vivo. We analyzed the subcellular localization of Flag-tagged HSD17B7 and HSD17B7E182* in HEI-OC1 cells. Since cholesterol synthesis mainly occurs in the endoplasmic reticulum (ER) [40, 41], HSD17B7 was expected to localize to this compartment. Consistent with this expectation, immunostaining in HEI-OC1 cells showed that Flag-tagged HSD17B7 co-localized with the ER marker Calnexin. In contrast, HSD17B7E182* failed to localize to the ER and instead displayed a spot-like aggregate distribution (Figure 7A), a pattern further supported by fluorescence intensity profile analyses (Figure 7B and C).

Subcellular localization of wild-type and p.E182* mutation HSD17B7.

(A) Immunostaining shows the subcellular localization of Flag-HSD17B7 and Flag-HSD17B7E182* in HEI-OC1 cells. Calnexin was used as an endoplasmic reticulum (ER) marker. White arrows indicate the line-scan positions used for intensity profile analysis in (B) and (C). Scale bars, 20 μm. (B, C) Line-scan intensity profiles showing the spatial distribution of HSD17B7 (B) or HSD17B7E182* (C) relative to the ER marker Calnexin along the indicated axes in (A). (D) Representative immunofluorescence image of a HEI-OC1 cell co-expressing HSD17B7 and HSD17B7E182*. White arrows indicate the positions used for line-scan analysis. Scale bars, 20 μm. (E) Line-scan intensity profiles showing the intracellular distribution of HSD17B7 and HSD17B7E182* along the indicated axis in (D).

Given that the patient is heterozygous for the variant, we next examined whether HSD17B7E182* interferes with HSD17B7 subcellular localization. Co-expression of Flag-HSD17B7 and HSD17B7E182*-Myc in HEI-OC1 cells showed that HSD17B7 retained its ER localization, indicating that the truncated protein does not significantly alter the subcellular distribution of the wild-type protein (Figure 7D and E).

To further investigate the function of HSD17B7E182* expression in vivo, we overexpressed human HSD17B7 or HSD17B7E182* in zebrafish HCs. The FM4-64 uptake assay revealed that the overexpression of HSD17B7 did not significantly affect FM4-64 fluorescence intensity in LLHCs compared to the control. In contrast, the expression of HSD17B7E182* led to a significant reduction of fluorescence intensity (Figure 8A and B). Consistently, startle response assays showed that HSD17B7 overexpression showed comparable movement trajectories, distance, and velocity in response to auditory stimulation, whereas larvae expressing HSD17B7E182* reduced startle-associated behaviors (Figure 8C-E). These data suggest that expression of HSD17B7E182* negatively impacts auditory-related behavior in vivo.

Expression of HSD17B7E182* impaired mechanotransduction activity, auditory-related behavior, and cholesterol organization.

(A) Representative images of lateral line hair cells (LLHCs, green) and FM4-64–labeled functional hair cells (red) in single neuromasts of Tg(Brn3c:mGFP) larvae at 5 dpf following injection of HSD17B7 or HSD17B7E182* mRNA. The white dashed indicates LLHCs. Scale bars, 20 μm. (B) Quantification of the FM4-64 relative intensity per LLHCs (n=25). One-way ANOVA followed by Tukey’s multiple comparisons, *P < 0.05, ****P < 0.0001, ns, non-significant (p > 0.05), mean ± SD. (C-E) Moving traces and quantification of 20 embryos were filmed in 150 frames before and after stimulation (9 dB re. 1 m·s⁻², 60 Hz tone burst). Scale bars, 10 mm. One-way ANOVA followed by Tukey’s multiple comparisons, *P < 0.05, **P < 0.01, ns, non-significant (p > 0.05), mean ± SEM. (F) Immunostaining of the cholesterol probe D4H in HEI-OC1 cells transfected with pCMV-Flag, pCMV-Flag-HSD17B7, and pCMV-Flag-HSD17B7E182* constructs, respectively. White dashed outlines indicate transfected cells; arrows denote line-scan positions used for intensity profile analysis. Scale bars, 20 μm. (G–I) Line-scan intensity profiles showing the spatial distribution of Flag, HSD17B7, or HSD17B7E182* relative to D4H signal along the indicated axes in (F). (J) Representative images of crista and lateral line hair cells in Tg(myo6b: D4H-mCherry) larvae following expression of myo6b-driven HSD17B7-EGFP or HSD17B7E182*-EGFP at 4 dpf. White dashed outlines indicate HCs. Scale bars, 10 μm.

Given our findings that Hsd17b7 regulates cholesterol homeostasis (Figure 4) and previous research that proper cholesterol distribution is critical for HC function [37], we next examined whether the HSD17B7E182* truncation disrupts intracellular cholesterol organization. HEI-OC1 cells were co-transfected with cholesterol probe D4H-mCherry together with pCMV-Flag, pCMV-Flag-HSD17B7, or pCMV-Flag-HSD17B7E182*. While cholesterol distribution appeared as a diffuse pattern around the nucleus in the control and HSD17B7-expressing cells, HSD17B7E182* resulted in a spot-like aggregation of D4H-mCherry (Figure 8F). Line-scan intensity analysis further confirmed the aberrant colocalization of HSD17B7E182* with cholesterol-enriched compartments (Figure 8G-I).

To validate these observations in vivo, HSD17B7 or HSD17B7E182* was expressed in zebrafish HCs by injecting pDest-myo6b-HSD17B7-EGFP or pDest-myo6b-HSD17B7E182*-EGFP plasmids into the Tg(myo6b: D4H-mCherry) transgenic line (Figure 8J). In HCs expressing HSD17B7, the protein exhibited a punctate intracellular distribution and did not noticeably alter the D4H-mCherry cholesterol signal compared with neighboring non-expressing hair cells. In contrast, HSD17B7E182* showed a diffuse localization throughout the cytoplasm and nucleus, accompanied by a marked reduction or near-complete loss of the D4H-mCherry signal in the stereocilia position. These data suggest that HSD17B7E182* with aberrant subcellular localization may bind cholesterol and alter its intracellular distribution. Altogether, these results demonstrate that HSD17B7E182* has a negative effect by altering cholesterol distribution in HCs, thereby compromising MET function and impaired startle responses.

HSD17B7E182* truncation disrupted the interaction with RER1

To further explore the pathological consequences of HSD17B7E182*, we compared the interaction landscapes of wild-type and mutant in HEI-OC1 cells (Figure 9A). Proteomic analysis of co-immunoprecipitated complexes revealed distinct interaction profiles between HSD17B7 and HSD17B7E182* (Figure 9A–C; Figure S7). Proteins uniquely associated with HSD17B7 were strongly enriched for cholesterol metabolic processes, whereas those preferentially associated with HSD17B7E182* were enriched for chromatin remodeling–related pathways.

HSD17B7 interacts with RER1 to ensure ER localization.

(A) Co-immunoprecipitation of Flag-tagged HSD17B7 and HSD17B7E182* from HEI-OC1 cells transfected with pCMV-Flag, pCMV-Flag-HSD17B7, or pCMV-Flag-HSD17B7E182*. Cell lysates were immunoprecipitated using anti-Flag beads. Tubulin and IgG light chains served as loading and immunoprecipitation controls, respectively. (B) Venn diagram showing the overlap and uniqueness of interacting proteins identified for HSD17B7 (red) and HSD17B7E182* (blue) by LC–MS/MS analysis. (C) Gene Ontology (GO) Biological Process enrichment analysis of proteins specifically associated with HSD17B7 or HSD17B7E182*. (D) GO Cellular Component enrichment analysis of HSD17B7- and HSD17B7E182*-specific interacting proteins. (E) Top 20 ER-localized proteins (ranked by iBAQ intensity) specifically associated with HSD17B7. (F) In vitro GST pull-down assays demonstrating direct interaction between HSD17B7 and RER1. The upper panel shows Coomassie blue staining of purified GST and GST-HSD17B7 proteins; the lower panel shows immunoblot detection of RER1-His. (G) Immunofluorescence staining of RER1-Myc in HEI-OC1 cells co-transfected with Flag-HSD17B7 or Flag-HSD17B7E182*. White arrows indicate positions used for fluorescence intensity profiling. Scale bars, 20 μm. (H, I) Fluorescence intensity profiles showing co-localization of RER1 with HSD17B7 (H) but not with HSD17B7E182* (I).

Consistent with their subcellular localizations, cellular component analysis showed that HSD17B7-interacting proteins were predominantly enriched in the endoplasmic reticulum (ER), whereas HSD17B7E182*-associated proteins were enriched in the nucleus and cytoplasm (Figure 9D), in agreement with immunostaining results (Figure 7A–C). Among ER-associated interactors, RER1 emerged as the top HSD17B7-specific binding partner (Figure 9E). RER1 is known to mediate ER retention of membrane-associated proteins [42, 43]. Direct binding between HSD17B7 and RER1 was confirmed by in vitro binding assays, whereas the HSD17B7E182* failed to interact with RER1 (Figure 9F). Immunofluorescence analysis further demonstrated robust co-localization of RER1 with HSD17B7 but not with HSD17B7E182* (Figure 9G–I), indicating that the p.E182* mutation disrupts RER1-mediated ER retention. Given that HSD17B7 is normally localized to the ER, loss of RER1 interaction provides a mechanistic explanation for the aberrant subcellular distribution of the mutant protein.

In addition to altered localization, the p.E182* mutation markedly reduced HSD17B7 expression levels. Compared with HSD17B7, the mutant protein exhibited substantially decreased abundance, accompanied by a significant reduction in mRNA levels (Figure S8A– C). mRNA stability assays revealed a shortened half-life of HSD17B7E182* transcripts relative to wild-type (Figure S8D and E), indicating that the nonsense mutation compromises transcript stability and consequently reduces protein expression.

Discussion

This study reveals that HSD17B7 is enriched in sensory hair cells in zebrafish and mice. Using the zebrafish model, we show that loss of hsd17b7 markedly reduces cholesterol levels in hair cells, leading to impaired MET function and abnormal hearing behaviors. These findings identify HSD17B7 as a previously unrecognized regulator of hair cell physiology and a candidate gene for sensory hearing loss. In addition to the animal models, we identified a previously undescribed heterozygous nonsense variant in HSD17B7 (c.544G>T, p.E182*) in a patient with sporadic deafness. Functional analyses revealed that the residual HSD17B7E182* protein alters subcellular localization and disrupts normal intracellular cholesterol distribution. Moreover, the mutation reduces mRNA half-life, decreases mRNA abundance, and significantly lowers protein expression. Collectively, these results suggest that the heterozygous c.544G>T (p.E182*) variant contributes to auditory dysfunction through potential pathogenic mechanisms: haploinsufficiency caused by reduced HSD17B7 expression and functional impairment due to altered cholesterol distribution. These mechanistic insights highlight the critical role of cholesterol homeostasis in MET and auditory-related function.

HSD17B7 was first identified as a prolactin receptor-associated protein in rats [44] and was proposed to be involved in estradiol biosynthesis [45, 46]. Subsequent biochemical studies demonstrated that HSD17B7 participated in cholesterol synthesis by catalyzing the reduction of zymosterone to zymosterol [47–49]. Consistent with this role, both in vivo and in vitro studies have shown that loss of HSD17B7 reduces cholesterol levels, whereas its overexpression increases cholesterol abundance. Previous studies have shown that HSD17B7 is expressed in the liver, heart, brain, eye, and ear [25, 26, 50], and scRNA-seq datasets have suggested its expression in mouse vestibular HCs [27]. Although the role of the HSD17B7 gene in ovarian and breast cancer has been extensively explored in the literature [51–55], the role of the HSD17B7 gene in auditory function remains unclear. Our results fill this gap by demonstrating that HSD17B7 is enriched in HCs, as supported by whole-mount in situ hybridization, immunostaining, and scRNA-seq analyses from our study and others [28, 31–33]. Given that loss of hsd17b7 causes abnormal auditory-related behaviors, these data establish HSD17B7 as a conserved and essential regulator of hair cell function.

Cholesterol is a fundamental component of biological membranes, contributing to both their structural integrity and functional properties [56, 57]. Previous studies have shown that either excessive or insufficient cholesterol levels, as well as abnormal cholesterol distribution, are detrimental to the auditory system [7, 10, 11, 36, 37, 58, 59]. However, previous studies have largely focused on systemic or membrane cholesterol alterations and have not addressed the contribution of hair cell–intrinsic cholesterol biosynthesis. Smith–Lemli–Opitz syndrome (SLOS), caused by mutations in DHCR7, exemplifies how impaired cholesterol synthesis can lead to multisystem defects, including sensorineural hearing loss, although auditory function in these patients has not been extensively investigated [60]. Our study revealed that HSD17B7, a cholesterol biosynthetic enzyme, is enriched in sensory hair cells and plays a crucial role in maintaining cholesterol levels required for auditory-related function. It should be noted that some localization and cholesterol distribution analyses were performed using C-terminal EGFP-tagged HSD17B7 constructs, which may influence protein behavior. However, the central conclusions of this study are supported by complementary loss-of-function models, rescue experiments using untagged mRNA, and independent cholesterol perturbation assays, reducing reliance on any single overexpression or fusion-based approach.

Here, we found that loss of Hsd17b7 or expression of the truncated HSD17B7E182* variant reduced cholesterol abundance and disrupted its intracellular distribution, resulting in defective MET function and auditory-related impairment. In contrast, overexpression of wild-type HSD17B7 elevated cholesterol levels but did not impair auditory-related behavior, suggesting that hair cell function requires a minimum critical cholesterol threshold rather than being sensitive to moderate cholesterol excess. Freeze-fracture studies have shown that the stereociliary membrane is densely enriched in cholesterol [61], supporting the notion that cholesterol is a key determinant of hair bundle membrane properties. Cholesterol is known to stiffen biological membranes [62], a biophysical feature that may influence MET channel gating [63, 64]. Consistent with this idea, cryo-EM studies have revealed conserved protein– lipid interactions within the TMC1/TMC2-containing MET channel complex [65, 66]. Together with our findings, these observations support a model in which HSD17B7 regulates MET function by maintaining appropriate cholesterol abundance within hair cells. How cholesterol perturbations caused by HSD17B7 deficiency or mutation mechanistically alter MET channel activity remains an important question for future studies.

Dominant non-syndromic hearing loss is a common form of hereditary deafness [67–74] [75–78], and de novo mutations are frequently observed in patients with a negative family history. In the present study, both parents of the proband exhibited normal hearing and declined genetic testing, suggesting that the HSD17B7 (c.544G>T, p.E182*) variant may represent a de novo dominant mutation. This interpretation is consistent with previous reports showing that homozygous deletion of Hsd17b7 in mice is embryonic lethal, whereas heterozygous phenotypes have not been systematically characterized [79]. Notably, the mutant mRNA failed to rescue auditory defects in hsd17b7 mutants, and overexpression of HSD17B7E182* in wild-type animals impaired startle response. Although additional clinical cases will be required to confirm pathogenicity, our functional analyses provide experimental support for a negative effect of the HSD17B7E182* variant. Together with the marked reduction in mRNA and protein stability, these findings suggest that the HSD17B7E182* variant primarily acts through loss of function, with the possibility that aberrant intracellular localization and misdistribution of cholesterol further exacerbate HC dysfunction.

In summary, our study identifies HSD17B7 as a critical regulator of cholesterol synthesis in sensory hair cells and as an essential factor in normal MET and sound-evoked sensory responses. Additionally, we identified a sporadic nonsense mutation in the HSD17B7 gene in patients with deafness and provided a mechanistic explanation. These findings uncover a previously unrecognized link between cholesterol biosynthesis and HC MET function and provide a molecular framework for exploring cholesterol-targeted therapeutic strategies for hearing loss.

Materials and Methods

Mouse and zebrafish husbandry

C57BL/6J mice were purchased from the Laboratory Animal Center of Nantong University and maintained in a barrier facility under a 12h light/12h dark cycle at 25℃. The day of vaginal plug observation was embryonic day 0.5 (E0.5), and the day of birth was defined as postnatal day 0 (P0). Zebrafish (Danio rerio) were raised and maintained at 28.5°C. Wild- type AB strain and the transgenic line Tg(Brn3c: mGFP) were used in the study described in our previous study [80]. The transgenic line Tg(myo6b: D4H-mCherry) was generated in this work. The embryonic stage is defined as described in the literature [81]. Embryos were collected following natural spawns and cultured in E3 medium (5 mM NaCl, 0.17 mM KCl, 0.33 mM CaCl2, and 0.33 mM MgSO4, pH 7.2). After 20 hours post-fertilization (hpf), embryos were incubated in E3 medium containing 0.2 mM phenylthiourea (PTU) to prevent pigmentation for in situ hybridization and imaging. All animal experiments were performed under the guidelines of the Institutional Animal Care and Use Committee of Nantong University.

Cell culture

The mouse auditory cell line HEI-OC1 was obtained from Prof. Renjie Chai and authenticated by short tandem repeat (STR) profiling (Shanghai Biowing Applied Biotechnology Co., Ltd). The cell lines were routinely tested negative for mycoplasma using the MycoBlue Mycoplasma Detection Kit (Vazyme Biotech, #D101-01), and cell aliquots from early passages were used. HEI-OC1 cells were cultured in DMEM (Wisent, #319-005-CL) medium supplemented with 10% FBS (Sigma-Aldrich, #F8318) in an incubator with 10% CO2 at 33℃.

HEK293T (sex unknown) cells were obtained from the Shanghai Institute of Biochemistry and Cell Biology (SIBC) and authenticated by short tandem repeat (STR) profiling (Shanghai Biowing Applied Biotechnology Co., Ltd). The cell lines were routinely tested negative for mycoplasma using the MycoBlue Mycoplasma Detection Kit (Vazyme Biotech, #D101-01), and cell aliquots from early passages were used. Cells were cultured in DMEM (Wisent, #319-005-CL) medium supplemented with 10% FBS (Sigma-Aldrich, #F8318) in an incubator with 5% CO2 at 37℃.

Antibodies

Primary antibodies used in this study included: anti-MYO7A monoclonal antibody (1:150, Developmental Studies Hybridoma Bank, #138-1); anti-HSD17B7 polyclonal antibody (1:150-1:1000, proteintech, #14854-1-AP); anti-Parvalbumin monoclonal antibody (1:500, sigma, #sab4200545); anti-MYC tag polyclonal antibody (1:200, proteintech, #16286-1-AP); anti-His antibody (1:2500, proteintech, # 66005-1-Ig); anti-DYKDDDDK tag mouse monoclonal antibody (1:200-1:5000, proteintech, #66008-4-Ig); anti-calnexin polyclonal antiboy (1:100, proteintech, #10427-2-AP); tubulin monoclonal antibody (1:5000, proteintech, #66031-1-Ig); goat anti-rabbit IgG(H+L) HRP (1:5000, MULTI SCIENCES, #GAR007); goat anti-mouse IgG(H+L) HRP (1:5000, MULTI SCIENCES, #GAM007); FITC goat anti-mouse IgG (H+L) (1:200, ABclonal, #AS001); Cy3 goat anti-rabbit IgG (H+L) (1:200, ABclonal, #AS007); anti-Digoxigenin-AP Fab fragments (1:1000, Roche, #11093274910).

RNA Isolation, Reverse Transcription (RT), and Quantitative Real-time PCR (RT-qPCR)

Total RNAs were extracted using TRIzol reagent (Invitrogen, #15596026) and treated with DNase I (Vazyme, #EN401) to remove genomic DNA contamination. RNA concentration was measured using a NanoDrop ND-2000 (Thermo Fisher Scientific, USA), and integrity was verified by agarose gel electrophoresis.

The cDNA was synthesized by using HiScript III RT SuperMix (Vazyme, #R323-01) following the manufacturer’s directions. Subsequently, quantitative PCR was performed using the ChamQ SYBR qPCR Master Mix (Vazyme, #Q341-02) with the specified primers (HSD17B7 primer, F: 5’-GACAAGCTTGGATCCATGCGAA -3’, R: 5’-ACCTGGACAA TGGTGACCTC -3’; Gapdh primer, F: 5’-CACAGTCAAGGCCGAGAATGGGAAG-3’, R: 5’-GTGGTTCACACCCATCACAAACATG) with a final volume of 20 μL under the following condition: 15 min at 50°C, 5 min at 95°C, and then 40 cycles at 95°C for 15 s and 60°C for 30 s. Relative expression levels for the hsd17b7 gene were calculated using the 2-△△CT method, normalized to the gapdh. All reactions were repeated in triplicate for each sample. The results were analyzed using the GraphPad Prism software (version 9.4.0).

Whole-mount in situ hybridization

The whole-mount in situ hybridization (WISH) of zebrafish was performed according to the following standard procedures. A 403 bp cDNA fragment of the zebrafish hsd17b7 gene or a 472 bp cDNA fragment of the zebrafish tmc1 gene was amplified via PCR using designed primers (hsd17b7-F: 5’-GACGTCCTCCAGTAATGCCC-3’, hsd17b7-R: 5’-CATCTTGCTTGGT CGGGTGT-3’; tmc1-F: 5’-TTGGGCAGTGATGTGCTGTA-3’, tmc1-R: 5’-GATGCTGTTT CTGCGTTGCT-3’) and was cloned into the pGEM-T-easy vector. After linearization of the pGEM-T-easy vector, inserting the hsd17b7 or tmc1 fragment, the DIG RNA Labeling Kit (SP6) (Roche, #11175025910) was used to prepare digoxigenin-labeled hsd17b7 or tmc1 antisense mRNA probes through transcription in vitro. Subsequently, embryos at different developmental stages were hybridized with an hsd17b7 or tmc1 mRNA probe overnight after a series of treatments, including fixation in 4% paraformaldehyde in PBS, digestion in proteinase K (Roche, #3115879001), and incubation with a pre-hybridized mix. Finally, the alkaline phosphatase (AP)-conjugated antibody against digoxigenin (Roche, #11093274910) and the AP-substrate NBT/BCIP solution (Roche, #11681451001) were used to detect the hsd17b7 or tmc1 expression. The WISH images of zebrafish were acquired using a stereomicroscope (Olympus, MVX10, Japan).

Plasmids construction

The HSD17B7 and RER1 full-length cDNAs were PCR amplified from the HEK293T cDNA and cloned into the pGEX-TEV and pET-23b vectors, respectively. HSD17B7E182* was generated by site-directed mutagenesis. For cell transfection, the cDNA of HSD17B7 and HSD17B7E182* were subcloned into the pCMV-Flag vector, respectively. The cDNA of RER1 and HSD17B7E182* were subcloned into the pcDNA3.1-Myc-His A vector, respectively. eGFP was subcloned into pCMV-Flag-HSD17B7 and pCMV-Flag-HSD17B7E182*, respectively. pCS2-D4H-mCherry was obtained from G. Peng’s lab. For plasmid injection of zebrafish, the myo6b promoter was subcloned into the pDestTol2 vector, followed by D4H-mCherry subcloned into the pDestTol2 vector. The cDNAs encoding the human protein of HSD17B7-eGFP and HSD17B7E182*-eGFP were inserted into the p-mTol2-myo6b vector, then linking P2A-D4H-mCherry by overlap extension. For mRNA injection of zebrafish, the cDNAs encoding the human protein of HSD17B7 and HSD17B7E182* were inserted into the pCS2+, and likewise, the cDNAs encoding the zebrafish full-length protein of Hsd17b7 were inserted into the pCS2+, which were linearized using NotI and then transcribed in vitro to generate mRNA. All constructs were verified by DNA sequencing.

Cell transfection

HEI-OC1 cells were transfected with plasmids using an X-treme GENE HP DNA transfection reagent (Roche, #6366236001) according to the manufacturer’s protocol. After 36 hours or 48 hours of transfection, the cells were fixed or collected and subjected to further analyses, including western blots, Co-IP, IF, or RT-qPCR.

HEI-OC1 cells were transfected with small interfering RNA. The mouse Hsd17b7 siRNA oligos (F: 5’-GGAGGUGUUUGAAACCAAUTT-3’, R: 5’-AUUGGUUUCAAACACCUCCTT-3’) were synthesized from GenePharma (Shanghai, China). Cells were transfected with siRNA oligos to knock down mHSD7B7 using Lipofectamine RNAiMAX (Invitrogen, #13778-150) according to the manufacturer’s protocols. Approximately 1×105 cells per well were placed in six-well plates. When the cell density reached 70-80%, 4 μL of control siRNA or mHsd17b7 siRNA was added to 150 μL of opti-MEM (Gibco, #31985070), then 5 μL Lipofectamine RANiMAX was added, incubated at room temperature for 5 min, and then added to the cell culture medium. After 48 hours of transfection, the transfected cells were harvested for Western blots or cholesterol assay.

mRNA stability assay

HEI-OC1 cells were transfected with plasmids using an X-treme GENE HP DNA transfection reagent. After 24 hours, the cells were treated with 10 µM actinomycin D (MCE, #HY-17559) for 0, 2, 4, 6, and 8 hours. Then, the collected cells and isolated RNA were used for RT-qPCR [82].

Immunostaining and image acquisition

Cultured cells grown on coverslips, after 36 hours of transfection, were fixed with 4% paraformaldehyde in PBS for 30 min at room temperature and washed 3 times with PBS, then permeabilized in 0.3% PBST (0.3% Triton X-100 in PBS) for 20 min. After three times washes with PBS, the cells were blocked with 10% FBS in PBS, followed by incubation with anti-primary antibodies overnight at 4°C, and then incubated with fluorescent dye-labelled secondary antibodies for 2 hours at room temperature before mounting using the mounting solution with DAPI (1:500, SouthernBiotech, #011-20).

The dissected mouse cochleae were fixed in 4% paraformaldehyde in PBS for 1 hour at room temperature. After 3 times washes with 0.1% PBST, the cochleae were briefly blocked with a blocking medium (PBS containing 1% Triton X-100 and 10% heat-inactivated donkey serum, pH 7.2) for 1 hour at room temperature, followed by incubation with the anti-primary antibody overnight at 4°C. The samples were then washed in 0.1% PBST and incubated with fluorescent dye-labelled secondary antibodies for 2 hours at room temperature, before mounting in mounting solution containing DAPI (1:500, SouthernBiotech, #011-20).

The immunostaining images of stained HEI-OC1 cells and cochlea were acquired by a Nikon confocal microscope with NIS-Elements software. For zebrafish live imaging, the larvae were anesthetized with tricaine MS-222 (Sigma, #A5040) and mounted in 0.6% low-melting agarose (Invitrogen, #16520050) with a lateral view. All reconstructed three-dimensional images and contrast adjustments were processed using Imaris X64 (version 9.0.1).

FM4-64 labeling

To investigate the basal activity of HCs, the FM4-64 (1:500, Invitrogen, #T13320) vital dye was used to specifically label functional HCs in the neuromasts. The staining procedures were carried out as described previously [29, 30, 83, 84]. The free-swimming larvae were incubated in 3 μM FM4-64 vital dye for 15 s at room temperature in the dark. Afterward, the fish were rinsed 3 times using a PTU medium and imaged. The images of stained hair cells were acquired by a Nikon confocal microscope with NIS-Elements software.

Measurement of total cholesterol

HEI-OC1 cells were homogenized in chloroform/methanol (2:1) to extract lipids and then centrifuged at 20,000 × g for 10 min. The organic phase was harvested and dried using nitrogen flow. Total cellular cholesterol was quantified using an Amplex Red cholesterol assay kit (Invitrogen, #A12216) through a multilabel reader (PerkinElmer, USA) according to the manufacturer’s instructions.

Quantification of the relative fluorescence intensity of FM4-64 and D4H

Quantitative analysis of FM4-64 uptake in zebrafish lateral line hair cells and D4H-mCherry fluorescence in HEI-OC1 cells was performed at the single-cell level. For zebrafish experiments, hair cells were identified based on Tg(Brn3c:mGFP). Individual hair cells within each neuromast were manually segmented using ImageJ, and the mean FM4-64 fluorescence intensity per hair cell was measured. Background fluorescence was determined from adjacent cell-free regions and subtracted from each measurement. For each larva, the average FM4-64 intensity was calculated by averaging values from all analyzed hair cells.

For D4H-mCherry cholesterol sensor analysis in HEI-OC1 cells, individual cells were segmented based on cell morphology, and the mean intracellular D4H-mCherry fluorescence intensity per cell was quantified. Background fluorescence was measured from cell-free regions within the same field of view and subtracted to normalize for imaging variability. Fluorescence intensities were then averaged across cells for each condition and used for statistical analysis. All fluorescence measurements were performed on raw images acquired under identical imaging settings across experimental groups.

Morpholino and mRNA injections

For inhibiting the expression of hsd17b7, hsd17b7-specific splicing-blocking morpholinos were designed and procured from Gene Tools, Inc., and the precise sequence was (5’-TGCAAACAGGTAACAAAACTGTGTG-3’). The morpholino powder was dissolved in RNase-free water to prepare the working solution at a final concentration of 0.3 mM for subsequent operations. About 2 nL dose of morpholino work solution was microinjected into zebrafish embryos at the one-cell stage. To assess morpholino efficiency, embryos injected with morpholino were collected, and RNA was extracted and reverse-transcribed into cDNA. The designed primers flanking on exon 1 and exon 2 (F: 5’-TACACAGGCAAACGTTAGAAGC-3’, R: 5’-CTTGAACTCCTCTGCACCCTT-3’) were used to amplify the fragment containing the mis-splicing target site, which was located at the connection of exon 1 and intron 1. For rescue experiments, exogenous mRNA was first transcribed in vitro. Briefly, the designed primers (zebrafish hsd17b7 mRNA primer, F: 5’-CGCGGATCCATGAAGAAAGTAGTTTTGGT-3’, R: 5’-CCGGAATT CTCACATTCCATTTCTTTCTT-3’; human HSD17B7 mRNA primer, F: 5’-CGGGATCCATGCGAAAGGTGGTTTTGATC-3’, R: 5’-GCTCTAGATTATAGGCATGA GCCACTGA-3’; human HSD17B7E182* mRNA primer, F: 5’-CGGGATCCATGCGAAAG GTGGTTTTGATC-3’, R: 5’-GCTCTAGACTAGAGGCTGAAATTAGATT-3’) were used to amplify target DNA containing the coding sequence. Subsequently, the pCS2 vector inserted into the amplified fragment was linearized as a template for mRNA transcription using the SP6 mMESSAGE mMACHINE Kit (Invitrogen, #AM1340). After purifying using the RNeasy Mini Kit (Qiagen, #74104), mRNA with a 70 ng/μL concentration was co-injected into one-cell stage embryos with morpholino for rescue experiments.

sgRNA/ Cas9 mRNA synthesis and injection

sgRNA was designed against the hsd17b7 gene (ENSDARG00000088140) using the CRISPR design tool (https://www.crisprscan.org). The sgRNA specifically targeting exon 3 of hsd17b7 (5’-GGGAATCATGCCTAATCCCA-3’) was first synthesized as follows. The GenCrispr sgRNA synthesis kit is used to generate a gRNA DNA template with a T7 promoter and to synthesize gRNA via in vitro transcription (Genscript Biotech Co., Ltd., Nanjing, China). The pXT7-zCas9 plasmid was linearized using XbaI and then transcribed in vitro to generate Cas9 mRNA using the mMESSAGE mMACHINE T7 kit (Invitrogen, #AM1344). At the one- cell stage, embryos were injected with a 2-3 µL solution containing 300 ng/µL Cas9 mRNA and 150 ng/µL sgRNA. The injected embryos were then grown in an E3 medium. At 24 hpf, genomic DNA was extracted from injected embryos, and potential CRISPR-induced mutations were identified by PCR. Primers used for identification were designed around the hsd17b7 sgRNA target sites (F: 5’-AAAAACTTATTTTATTCCAGCCCAA-3’, R: 5’-TTTCCAT GCAGCACTATCAAACAATT-3’).

Startle Response Test

The acoustic startle reflex was performed as described previously [34, 85]. Briefly, A plastic plate attached to a mini vibrator was used to place 20 normal larvae at 5 days post-fertilization (dpf), while an infrared digital video tracking system was used to monitor their swimming behavior. 60 Hz tone bursts at two sound levels of 9 dB re. 1 m·s −2 were applied to the amplifier to drive the vibrator. Acoustic vibration stimuli lasting 30 ms, with an inter-stimulus interval of 180 s, were set and applied. Each sound vibration stimulus level was repeated 20 times, and the locomotion behavior of the larvae with C-shape motion in response to this stimulus was recorded. Finally, the movement’s typical parameters of mean distance and peak velocity were analyzed to assess the startle response of larvae to sound vibration stimuli.

In vitro binding assay

GST and GST-HSD17B7 fusion proteins were immobilized on GSH resins and incubated with purified His-RER1 proteins overnight at 4℃ and extensively washed with washing buffer (0.3% PBST). Bound proteins were separated using SDS-PAGE and visualized with Coomassie Blue staining and Western blots.

Western blots

Frozen cells were lysed in RIPA buffer (50 mM Tris-HCl, 150 mM NaCl, 1 mM EDTA, 1% sodium deoxycholate, 1% w/v protein inhibitor (Roche, #4693132001)) for 30 min and centrifuged at 12,000 g for 15 min at 4℃. Protein concentrations were determined using a BCA protein assay kit (Thermo Fisher, #23227).

The proteins were separated on SDS-PAGE gels and subsequently transferred to a PVDF membrane (Millipore, USA). The membranes were blocked in blocking buffer (1 × TBS containing 0.5% milk and 0.5% Tween 20) for about 45 min at room temperature, then incubated overnight with primary antibodies in TBS containing 4% BSA, 1% Tween20, and 0.05% NaN3. The next day, the PVDF membranes were incubated with the appropriate HRP-conjugated secondary goat antibodies for 2 hours at room temperature after extensive washing with a blocking buffer. The blots were detected using ECL TM Western Blotting Detection Reagents (GE, #RPN2106), and the acquired images were analyzed using ImageJ (version 1.8.0).

Co-IP, Mass spectrum, and Gene Ontology (GO) analysis

Cell lysates from pCMV-Flag, pCMV-Flag-HSD17B7, and pCMV-Flag-HSD17B7E182* transfected cells were incubated with 30 μL of beads conjugated to anti-Flag antibodies (Smart-Lifesciences, #SA042001), respectively, overnight at 4℃. The next day, Flag beads were washed 3 times with lysis buffer, and the immune complexes were eluted with 2× SDS sample buffer and subjected to three samples for SDS-PAGE and mass spectrum analysis.

For mass spectrum, the immunoprecipitated HSD17B7 and HSD17B7E182*-associated proteins were washed and dissolved in ammonium bicarbonate buffer (25 mM, pH 8.0), followed by trypsin digestion at 37℃ for 16 hours. The nano-LC-MS/MS experiments were performed with LTQ-Orbitrap MS (Thermo Fisher, USA) equipped with a Nano electrospray ion source.

For the Venn diagram and GO analysis, microarray analysis identified HSD17B7 and HSD17B7E182*-specific associated proteins by mass spectrum (peptide ≥ 1, unique peptide ≥ 1, P value < 0.05). Cellular component GO term enrichment analysis was performed using the web-based DAVID software [86]. Ontology networks were further investigated and visualized using R (version 4.4.1).

Public scRNA-seq datasets and expression profiling of hsd17b7

Zebrafish scRNA-seq data were downloaded from NCBI Gene Expression Omnibus (GSE221471) [28]. The basic procedure for single-cell integrated data analysis was performed using Seurat 4.0.1 [87]. Cell clusters corresponding to LLHCs, MHCs, CHCs, supporting cells, and mantle cells were selected based on previously published annotations. Feature plots, violin plots, and average expression analyses were used to examine the expression pattern of hsd17b7 across cell types. To assess dynamic changes in hsd17b7 expression during HC differentiation, trajectory analysis was performed using Monocle 3 (v1.0.1). Cells were ordered along pseudotime, with mantle cells as the root population, and gene-expression changes were visualized along the inferred developmental trajectory.

Mouse scRNA-seq data were downloaded from previous articles (GSE71982, GSE168901, GSE202920) [31–33]. The basic procedure for single-cell integrated data analysis and batch-effect correction was performed using Seurat 4.0.1 [87]. We selected four cell populations for re-analysis, including inner hair cells, outer hair cells, utricle hair cells, and crista hair cells.

Single-cell RNA sequencing of control and hsd17b7 mutant hair cells

To investigate genotype-dependent transcriptional remodeling in sensory hair cells, we performed single-cell RNA sequencing on fluorescence-activated cell sorting (FACS)-isolated hair cells from control and hsd17b7 mutant zebrafish.

Isolation of hair cells and single-cell RNA sequencing

mGFP-positive hair cells were isolated from control and hsd17b7 mutant zebrafish larvae by fluorescence-activated cell sorting (FACS). Dissociated cells were gated to exclude debris and doublets, and viable mGFP⁺ cells were collected for downstream analysis. Single-cell RNA sequencing libraries were generated using the 10× Genomics Chromium platform according to the manufacturer’s instructions. Libraries were sequenced on an Illumina platform to obtain paired-end reads. Raw sequencing data were processed using the Cell Ranger pipeline (10× Genomics) to generate gene expression matrices for each sample. Downstream analyses were performed in R using the Seurat package.

Data preprocessing, integration, and cell type annotation

Cells with low gene counts or high mitochondrial transcript proportions were excluded to remove low-quality cells. Gene expression matrices from control and hsd17b7 mutant samples were normalized and integrated using Harmony to correct for batch effects while preserving biological variation.

Dimensionality reduction was performed using principal component analysis (PCA), followed by Uniform Manifold Approximation and Projection (UMAP) for visualization. Clustering was conducted using a shared nearest neighbor (SNN) graph-based approach. Cell clusters were annotated based on established marker genes for hair cells and supporting cells. Hair cells were identified by robust expression of myo6b, while inner ear and lateral line hair cell subtypes were distinguished by lhfpl5a and lhfpl5b, respectively. Supporting cell populations were annotated using stm as a pan–supporting cell marker, with otogl and irg1l marking inner ear and lateral line supporting cells, respectively. Cell type annotations were validated using both violin plots and Nebulosa density visualizations

Pseudobulk differential expression analysis

To assess genotype-dependent transcriptional changes while accounting for cell-to-cell variability, pseudobulk differential expression analysis was performed for hair cells. Gene expression counts from individual cells were aggregated by genotype to generate pseudobulk profiles. Differential expression between control and hsd17b7 mutant HCs was assessed using a generalized linear model framework. Genes with an adjusted p-value < 0.05 were considered differentially expressed. Results were visualized using volcano plots, with upregulated and downregulated genes highlighted accordingly.

Gene Ontology and pathway enrichment analysis

Differentially expressed genes were subjected to Gene Ontology (GO) biological process enrichment analysis using curated annotation databases [88]. Enrichment analyses were performed separately for upregulated and downregulated gene sets. Significantly enriched GO terms were visualized, and representative genes contributing to each term were illustrated using Sankey diagrams to highlight functional relationships between gene sets and biological processes.

Gene set enrichment analysis (GSEA) was additionally performed to assess coordinated changes in cholesterol biosynthesis–associated pathways at the transcriptome-wide level. Ranked gene lists derived from pseudobulk differential expression results were used as input, and enrichment significance was assessed using normalized enrichment scores and adjusted p-values.

Module score analysis of MET- and cholesterol-related gene sets

Module scores were computed using the AddModuleScore function in Seurat, based on normalized expression values from the RNA assay. Module score distributions were examined separately in lateral line hair cells (LL_HCs) and inner ear hair cells (IE_HCs). Differences between control and hsd17b7 mutant samples were assessed using the Wilcoxon rank-sum test. Results were visualized using heatmaps, violin plots, and ridge plots. Ridge plots were generated using z-scored module scores to illustrate population-wide shifts in transcriptional states rather than the emergence of discrete subpopulations.

Visualization and statistical analysis

UMAP feature plots, heatmaps, violin plots, and ridge plots were generated using Seurat and associated R visualization packages. Nebulosa density plots were used to visualize spatial expression patterns of marker genes across low-dimensional embeddings. Statistical analyses were performed using R, and exact p values or significance levels are reported in the corresponding figure panels and legends.

Whole-exome sequencing and Sanger sequencing analyses

Whole-exome sequencing was performed in the proband Ⅱ 1. Genomic DNA samples were extracted from whole blood samples of the proband Ⅱ 1 (Figure 5A).

The exomes and flanking intronic regions from whole blood DNA samples were captured by Agilent SureSelect Human All Exon Kit (Agilent Technologies, USA). The captured DNA was sequenced on the Illumina HiSeq 4000 sequencing platform (Illumina, USA). Bioinformatics were aligned to the NCBI build37/hg19 assembly using the BWA (version 0.7.12) software. Each sample was covered to an average sequencing depth of at least 100×. SNPs and indels were identified using the GATK HaplotypeCaller software. Candidate pathogenic variants were defined as nonsense, missense, splice-site, and indel variants with allele frequencies of 0.001 or less in public variant databases (dbSNP, 1000 Genomes, ESP6500, nci60, GnomAD) and in disease databases (COSMIC, ClinVar, OMIM, GWAS). Genotypes distributed every 0.3 cM of genomic region were chosen for calculation of the logarithm of odds (LOD) scores using the Merlin v. 1.1.2 parametric linkage analysis package. The proband Ⅱ 1 was genotyped by Sanger sequencing analyses with designed primers (F: 5’-GTACTCTGATTGGTGACGGGTGAG-3’, R: 5’-GACAGTCATAGTTCATAGTTTATT-3’)

Quantification and statistical analysis

GraphPad Prism 9 (version 9.4.0) supported the whole statistical analysis. All data presented as mean ± SD or mean ± SEM. An unpaired two-tailed Student’s t-test was performed for two-group comparisons, while multiple comparisons were illustrated using one-way ANOVA. The p-value less than 0.05 (p ≤ 0.05) was considered significantly different. p ≤ 0.05, p ≤ 0.01, p ≤ 0.001, and p ≤ 0.0001 were symbolized with “*, **, ***, and ****”, respectively, and “ns” represented no significance, p > 0.05.